### 2020-09-30 paper/policar2020-opentsne.tex
Edit methods section

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

\documentclass[twocolumn]{bmcart}
%\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\usepackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

\usepackage{graphicx}
\usepackage{minted}
\usemintedstyle{vs}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\newcommand{\opentsne}{\textsf{openTSNE}}
\endlocaldefs

\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Software}

\title{openTSNE: a modular Python library for t-SNE dimensionality reduction and embedding}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={pavlin.policar@fri.uni-lj.si}   % email address
]{\inits{PGP}\fnm{Pavlin G.} \snm{Poli\v{c}ar}}
\author[
   addressref={aff1,aff2}
]{\inits{MS}\fnm{Martin} \snm{Stra\v{z}ar}}
\author[
   addressref={aff1,aff3}
]{\inits{BZ}\fnm{Bla\v{z}} \snm{Zupan}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%
  \orgname{Faculty of Computer and Information Science, University of Ljubljana},
  %\street{Ve\v{c}na Pot 113},
  \postcode{SI 1000}
  \city{Ljubljana},
  \cny{Slovenia}
}
\address[id=aff2]{%
  \orgname{Broad Institute of Harvard and MIT},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{MA 02142}
  \city{Cambridge},
  \cny{U.S.A.}
}
\address[id=aff3]{%
  \orgname{Department of Molecular and Human Genetics, Baylor College of Medicine},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{TX 77030}
  \city{Houston},
  \cny{U.S.A.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract}
Point-based visualisations of large, multi-dimensional data from molecular
	biology can reveal meaningful clusters. One of the most popular
	techniques to construct such visualisations is t-distributed stochastic
	neighbor embedding (t-SNE), for which a number of extensions have
	recently been proposed to address issues of scalability and the quality
	of the resulting visualisations. We introduce \opentsne, a modular
	Python library that implements the core t-SNE algorithm and its
	extensions. The library is orders of magnitude faster than existing
	popular implementations, including those from scikit-learn. Unique to
	\opentsne\ is also the mapping of new data to existing embeddings,
	which can surprisingly assist in solving batch effects.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{t-SNE}
\kwd{embedding}
\kwd{visualization}
\kwd{dimensionality reduction}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}

The abundance of high-dimensional data sets in molecular biology calls for
techniques for dimensionality reduction, in particular for methods that produce informative data visualizations. Popular approaches include 
principal component analysis (PCA),
multidimensional scaling, t-distributed stochastic neighbor embedding
(t-SNE)~\cite{maaten2008visualizing}, and uniform manifold approximation and
projections (UMAP)~\cite{2018arXivUMAP}. 
Among these, t-SNE lately received much attention as it can address high
volumes of data and reveal meaningful clustering structure. Increased
resolution and throughput of modern molecular assays has lead to frequent use
of t-SNE in diverse fields including, but not limited to, single-cell
transcriptomics(scRNA-seq,~\cite{macosko2015highly,cao2019single,tasic2018shared}),
human genetics~\cite{hirata2019genetic}, metagenomic
assembly~\cite{beaulaurier2018metagenomic}, spatial organization of microbial
communities~\cite{sheth2019spatial} and
metabolomics~\cite{tkachev2019differences}.

Reports on single-cell gene expression data, our running example,
often start with an overview of the cell landscape, where t-SNE embeds
high-dimensional expression profiles into a two-dimensional
space. Figs.~\ref{fig:macosko}.a and \ref{fig:macosko}.b are examples of two such
embeddings.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{macosko2015}
  \caption{\label{fig:macosko}The figure shows the progression of t-SNE
	embeddings following theoretical advances, which are all included in
	\opentsne. The data in \textbf{(a)} and \textbf{(b)} represent 44,808
	single-cell gene-expression profiles of mouse retinal cells from the
	study by Macosko \textit{et al.}~\cite{macosko2015highly}. Similarly,
	the data in \textbf{(c)} contains 27,499 expression profiles from mouse
	retinal cells from Shekhar \textit{et
	al.}~\cite{shekhar2016comprehensive}. \textbf{(a)} We construct a t-SNE
	embedding following the parameter choices from the original publication
	from Maaten \& Hinton~\cite{maaten2008visualizing}. The visualization
	shows no preservation of the global organization of clusters,
	resulting from random initialization and an affinity model focused on
	preserving local neighborhoods. \textbf{(b)} shows a modern t-SNE
	embedding, utilizing the latest theoretical advances and practical
	recommendations. The embedding is constructed using a multi-scale
	affinity model, preserving both short-range and long-range interactions
	between data points, and is initialized such that the global layout is
	meaningful. Unlike in \textbf{(a)}, the green and blue clusters
	representing different sub-types of amacrine and bipolar cells are now
	localized to the same regions of the space, indicating a higher level
	of similarity than to other cell types. \textbf{(c)} shows how existing
	t-SNE reference atlases can be used to embed new samples into existing
	embeddings. The positions of new data points correspond to cell types
	from the reference atlas.}
\end{figure*}

Despite its utility, t-SNE has often been criticized for its limited scalability, lack of global organization -- t-SNE emphasizes
local clusters that are arbitrarily scattered in the low-dimensional space --
and the absence of theoretically-founded methods to map new data into
existing embeddings~\cite{ding2018interpretable,becht2019dimensionality}. Most
of these shortcomings have recently been addressed. Linderman \textit{et al.}
sped-up the method through an interpolation-based approximation, achieving linear complexity in the number of
samples~\cite{linderman2019fast}. Kobak \& Berens proposed several techniques
to improve global positioning, including estimating similarities with a mixture
of Gaussian kernels~\cite{kobak2019art}. In our previous work, we presented a
principled approach for embedding new samples into existing
visualizations~\cite{policar2019embedding}.

\section*{Results}

We introduce \opentsne, a comprehensive Python library that implements t-SNE
and all its recently proposed extensions. Most notable among these are
\begin{enumerate}
\item the implementation of efficient approximation 
	schemes~\cite{van2014accelerating,linderman2019fast} allowing the embedding
	of millions of data points,
\item the addition of new samples into existing
	embeddings~\cite{policar2019embedding},
\item better initialization schemes~\cite{kobak2019umap} leading to more
	globally consistent layouts,
\item variable degrees of freedom~\cite{kobak2019heavy} allowing the inspection
	of data at different levels of resolution,
\item multi-scale similarity kernels~\cite{kobak2019art} which preserve small,
	well-defined clusters and uncover global relationships between clusters,
	and
\item improved defaults for learning rate and number of
	iterations~\cite{belkina2019automated} which produce embeddings at a low
	computational cost.
\end{enumerate}

The library is fully compatible with the Python data-science ecosystem (e.g.,
\textsf{numpy}, \textsf{scikit-learn}, \textsf{scanpy}) providing a familiar
and intuitive API to new users. Its modular design encourages
extensibility and experimentation with various settings and changes in the
analysis pipeline.

\opentsne\ is open-source and has become adopted by the wider community,
receeiving over 600 Github stars and averaging thousands of weekly downloads.
The source code is publicly available on Github at
\url{https://github.com/pavlin-policar/openTSNE}, and can be installed from
\textsf{PyPI} and \textsf{conda-forge} -- the two most widely adopted Python
package managers.

Accessibility is one of the core design principles of \opentsne. Our aim is to
make the latest theoretical advances in t-SNE as accessible as possible to as
many users as possible. As such, our API is based on the familiar
\textsf{scikit-learn} interface~\cite{sklearn_api}, which enables users a
smooth transition from existing Python libraries. For example,
Fig.~\ref{fig:tasic} depicts four different t-SNE visualizations generated by
\opentsne, each based on recent theoretical advances in the t-SNE algorithm. The code used
to generate these embeddings is shown below.

\begin{minted}[
fontsize=\footnotesize,
frame=lines,
framesep=2mm,
baselinestretch=1.1
]{python}
import openTSNE
from openTSNE.affinity import Multiscale
# a - standard t-SNE
tsne1 = openTSNE.TSNE(perplexity=30).fit(X)
# b - high perplexity  for globa structure
tsne2 = openTSNE.TSNE(perplexity=500).fit(X)
# c - multiscale kernel for local/global structure
aff = Multiscale(X, perplexities=[30, 500])
tsne3 = openTSNE.TSNE(affinities=aff).fit(X)
# d - decrease dof for higher resolution
tsne4 = openTSNE.TSNE(dof=0.6).fit(X)
\end{minted}

\noindent The code snippet takes as input a \textsf{numpy} array or
\textsf{scipy} sparse matrix \texttt{X} and produces four
\texttt{TSNEEmbedding} objects. These may subsequently be subject to further
optimization under different parameter settings or may be used to embed new
samples into the embedding landscape. Additional parameters are omitted for
clarity but can be found in accompanying notebooks publicly available at
\url{https://github.com/pavlin-policar/opentsne-paper}.

\section*{Discussion}

\subsection*{Uncovering Structure in High-Dimensional Data}

Dimensionality reduction techniques implicitly assume that high-dimensional
data lies on a lower-dimensional manifold, which can accurately be captured by
a smaller number of dimensions. However, there is no evidence that
every data set can accurately be described by using only two dimensions, and any such embedding will therefore inevitably lead to loss of
information. It is thus beneficial to examine multiple embeddings, each
of which provides a different perspective into topology and other
data characteristics.

\begin{figure*}[htbp]
	\includegraphics[width=\textwidth]{tasic2018}
	\caption{\label{fig:tasic}We use openTSNE to create four different
    visualizations of the Tasic \textit{et al.}~\cite{tasic2018shared} data,
    each providing a different perspective into the topology of the data.
    The data set contains 21,874 single-cells originating from the mouse
	neocortex. Cluster annotations and colors are taken from the original
	publication. Warm colors correspond to excitatory neurons, cool colors
	correspond to inhibitory neurons, and gray/brown colors correspond to
	non-neuronal cells. The standard t-SNE \textbf{(a)} emphasizes local
	structure while increasing perplexity \textbf{(b)} results in a more
	meaningful layout of the clusters. We can also combine the two
	perplexities by using a multiscale kernel affinity model \textbf{(c)}
	and obtain a good trade-off between global and local structure.
	Alternatively, we can inspect more fine-grained structure and reveal
	smaller clusters by using a more heavy-tailed kernel \textbf{(d)}.}
\end{figure*}

We illustrate this point by generating four embeddings of the same data on
single-cell gene expression in mouse brain~\cite{tasic2018shared}.  First,
Fig.~\ref{fig:tasic}.a shows an embedding using default t-SNE parameters. While
different clusters of excitatory and inhibitory neurons appear close to one
another, all clusters appear equidistant from their neighbors, and the overall
relations between groups are not obvious. On the other hand,
Fig.~\ref{fig:tasic}.b focuses on preserving larger neighborhoods of points,
resulting in a more globally consistent layout where relations between clusters
become more apparent. In this embedding, it is evident from the increased
white space between groups that there is one large class of excitatory neurons,
and two related classes of inhibitory neurons. Unfortunately, focusing on
preserving large neighborhoods often results in smaller clusters being absorbed
into larger ones. Alternatively, Fig.~\ref{fig:tasic}.c uses multi-scale
similarity kernels which aim to preserve both the global organization of
clusters as well as prevent smaller clusters from being absorbed into larger
ones. On the other hand, Fig.~\ref{fig:tasic}.d shows use the same parameter
values as Fig.~\ref{fig:tasic}.a, but at a finer level of resolution. This
reveals that many of the clusters are composed of other subgroups and may be
clustered further into cell subtypes.

When working with data containing millions of data points, standard t-SNE
embeddings often become unwieldy -- cluster boundaries become blurred, large
clusters absorb smaller ones, and relationships between clusters become
increasingly difficult to interpret. Fig.~\ref{fig:cao}.a contains expression
profiles of over two million single cells captured at different time points in
mouse development. The embedding indicates numerous clusters with some
transition between time points as indicated by the color-coding, but this is
largely difficult to interpret. Kobak \& Berens observed that increasing
attractive forces between similar data points controlled via the
\textit{exaggeration} parameter leads to more compact clusters, and
subsequently more informative visualizations~\cite{kobak2019art}. For instance,
\ref{fig:cao}.b doubles the default exaggeration  which begins to reveal the
overall structure of the data. Doubling the exaggeration again in
\ref{fig:cao}.c, we can now easily observe that the data is comprised of two
main groups of cells and eight somewhat smaller clusters. The visualization
also reveals a number of tiny clusters, possibly corresponding to
rare cell types.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{cao2019}
  \caption{\label{fig:cao}Increasing the exaggeration parameter produces more
  	compact clusters, which highlight the global organization of the clusters
  	and emphasizes continuities between cell states. The data set from Cao \textit{et
	al.}~\cite{cao2019single} contains expression profiles from 2,058,652
	single cells. The data were collected from mice embryos at different
	developmental time-points at daily intervals after 9.5 to 13.5 days.
	\textbf{(a)} shows the data is comprised of two main components -- the
	neural tube and mesenchymal cells -- as well as several other smaller
	clusters. Additionally, the colors indicate developmental progression
	with red indicating least-developed cells and blue indicating most
	developed cells. The overall developmental trajectory is most apparent
	with higher levels of exaggeration, showing red cells slowly
	transitioning into blue cells. Progressively easing the exaggeration
	factor uncovers finer clusters within the larger groups as shown in
	\textbf{(b)} with exaggeration 2 and subsequently in \textbf{(c)},
	where we show the standard t-SNE with no exaggeration. 32,011 putative
	doublets are excluded from the visualizations.}
\end{figure*}

Exaggeration can highlight transitions between cell-states in
developmental studies. Standard t-SNE often produces embeddings with
clearly defined, discrete clusters. We have shown that we can adjust the level
of granularity and resolution of the clusters with several parameters in Fig.~\ref{fig:tasic}. However,
discrete clusters are often undesired in developmental studies where cells are assumed to follow a continuous trajectory between cell states. To this end, other
embedding techniques such as UMAP and ForceAtlas2~\cite{jacomy2014forceatlas2} tend to better capture the continuity between cell
states. Recently, B{\"o}hm \textit{et al.}~\cite{bohm2020unifying} have shown that embeddings
produced by t-SNE with exaggeration values of 4 and $\sim30$ produce embeddings
which are markedly similar to UMAP and ForceAtlas2, respectively. For example,
in Fig.~\ref{fig:cao}.a, the developmental trajectory between different time
points is difficult to observe due to a large number of sprawled out clusters.
On the other hand, the development becomes easier to follow when the
exaggeration factor is increased from $1$ to $4$ (Fig.~\ref{fig:cao}.b-c).

\subsection*{Embedding New Samples}

Unlike other popular dimensionality reduction methods such as principal
component analysis or autoencoders, t-SNE is a non-parametric method and does
not define an explicit mapping to the embedding space. Therefore embeddings of
new data points need to be found through the use of optimization
techniques~\cite{policar2019embedding}. \opentsne\ is currently the only
publicly available library allowing users to add new samples to existing
embeddings.

Figs.~\ref{fig:macosko}.b and \ref{fig:macosko}.c demonstrate how we can use a
previously labeled single-cell data set and embed cells from a separate
experiment into the reference landscape. The reference data taken from Macosko~\textit{et al.}~\cite{macosko2015highly} contains gene expression profiles from
mouse retinal cells. By embedding the samples from a similar experiment on bipolar retinal cells by Shekhar~\textit{et al.}~\cite{shekhar2016comprehensive}, 
we can correctly map to the bipolar cell clusters in the reference.

Embedding single cells into existing reference atlases can also be useful for
cell type classification in cases where cell identities are not known in
advance. For instance, in Fig.~\ref{fig:transform}, we construct a reference
embedding using labeled data from Hochgerner~\textit{et
al.}~\cite{hochgerner2018conserved} containing gene-expression profiles of
cells from the mouse brain. The authors assign a cell type to each cell, and
we can verify their accuracy by visualizing well-established gene markers for the
major cell types. We then embed cells from Harris~\textit{et
al.}~\cite{harris2018classes}, where labels are provided only for neuronal
cells. We quickly identify other non-neuronal cell types, including
oligodendrocytes and astrocytes. We use marker genes to validate that new cells
were correctly mapped into the reference landscape.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{transform_hochgerner}
  \caption{\label{fig:transform}openTSNE enables the user to embed new samples
	into an existing, reference embedding. We construct a t-SNE embedding
	for the data from Hochgerner \textit{et
	al.}~\cite{hochgerner2018conserved} containing 24,185 developing,
	single cells from the mouse hippocampus. The data contains different
	types of neurons, supporting glia, and other vascular cells (upper
	left). Cell labels are taken from the original publications. The
	colors correspond to Fig.~\ref{fig:tasic}. We then embed new,
	hippocampal cells collected by Harris \textit{et
	al.}~\cite{harris2018classes} using the previous embedding as a
	reference. In their study, Harris \textit{et al.} collected 6,971
	single-cells and focused on identifying different types of inhibitory
	neurons. However, almost half of the collected cells are not neurons
	and were left uncharacterized. Inspecting where these cells are
	embedded in the reference embedding (bottom left) reveals that in
	addition to inhibitory neurons, the data contains several supporting
	glial cells as well as a small population of endothelial cells. We can
	verify the accuracy of our approach by inspecting marker genes for the
	major cell types in the reference (top row) and embedded samples
	(bottom row).}
\end{figure*}

These two examples demonstrate how \opentsne\ can be used to quickly gain
insight into newly-sequenced, single-cell data sets by utilizing existing cell
atlases. This approach is general and not limited to single-cell gene
expression, but can be used with any tabular data set regardless of field.

\subsection*{Versatility}

Versatility is another of  \opentsne s core design principles. Kobak \& Berens
recently introduced several recommendations and tricks to obtain better and
more meaningful t-SNE visualizations~\cite{kobak2019art}. These include the use
of multi-scale similarity kernels, perplexity annealing, and increasing
exaggeration when working with larger data sets. \opentsne\ provides a flexible
API, where these improvements can be incorporated in just a few lines of code.
Furthermore, \opentsne\ supports custom affinity models, enabling users to
construct t-SNE embeddings on non-tabular data, e.g. relational data, as the
only affinity-model requirement is a notion of similarity between data points,
i.e. the distance matrix.  Finally, our comprehensive callback system can be
utilized to monitor and adapt different stages of the optimization phase and
has been used to construct visually appealing animations of the t-SNE
optimization process.

\subsection*{Speed}

One of t-SNE's common criticisms is limited scalability to large data sets
containing up to millions of data points~\cite{becht2019dimensionality}. These
criticisms are largely based on two factors. Until quite recently, most popular
implementations of t-SNE were based on the Barnes-Hut approximation scheme
developed by van der Maaten in 2014~\cite{van2014accelerating} with asymptotic
complexity $\mathcal{O}(N \log N)$. More importantly, this criticism was
exacerbated further by the fact that the most widely-used implementation of
t-SNE came from \textsf{scikit-learn}, which exhibits long runtimes
performance compared to its C++ counterpart -- \textsf{MulticoreTSNE}. The
multi-threaded \textsf{MulticoreTSNE} implementation~\cite{Ulyanov2016} is
perfectly capable of creating t-SNE embeddings of millions of data points in a
matter of hours on widely-accessible, consumer-grade processors
(Fig.~\ref{fig:benchmarks}). However, recently, Linderman \textit{et al.}
developed a new approximation scheme -- FIt-SNE -- which further reduces the
asymptotic time complexity to $\mathcal{O}(N)$, enabling the embedding of large
data sets in a matter of minutes.

Fig.~\ref{fig:benchmarks} benchmarks four Python t-SNE implementations,
including those from \textsf{scikit-learn} (v0.23.1), \textsf{MulticoreTSNE}
(v0.1), \textsf{FIt-SNE} (v1.1.0), and \opentsne\ (v0.4.3). We perform
benchmarks on two processors, one representing usage on a personal computer and
one on a high-performance computing platform. The Intel(R) Core i7 is commonly
found in consumer-grade laptop computers, while Intel(R) Xeon(R) processors are
found in high-performance computing machines. Benchmarks were run for 1000 iteration with the original t-SNE parameters, as some implementations do not
allow their modification.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{benchmarks}
  \caption{\label{fig:benchmarks}We benchmark openTSNE (v0.4.3) against three
	popular open-source implementations from
	scikit-learn~\cite{pedregosa2011scikit} (v0.23.1),
	MulticoreTSNE~\cite{Ulyanov2016} (v0.1), and
	FIt-SNE~\cite{linderman2019fast} (v1.1.0). Benchmarks were run on a
	consumer-grade Intel Core i7-7700HQ processor found in laptop
	computers, and on a server-grade Intel Xeon E5-2650. To generate
	benchmark data sets of different sizes, we subsampled data from the 10X
	Genomics 1.3 million mouse brain data set five times, resulting in five
	different data sets for each size. In total, each implementation was
	run on 30 different data sets. Notice that openTSNE scales similarly to
	FIt-SNE, as they both use the same interpolation-based approximation
	scheme, while scikit-learn and MulticoreTSNE utilize the Barnes-Hut
	approximation.}
\end{figure*}

Fig.~\ref{fig:benchmarks} demonstrates that both \textsf{FIt-SNE} and
\opentsne\ exhibit better scaling than their Barnes-Hut
counterparts -- \textsf{scikit-learn} and \textsf{MulticoreTSNE}. Notice that
even though \textsf{FIt-SNE} and \opentsne\ both implement the same
approximation scheme, \opentsne\ tends to be slightly slower. \opentsne\ is
written in Python, which inevitably incurs some runtime overhead compared to
its C++ counterpart. However, Fig.~\ref{fig:benchmarks} shows the
runtime of \opentsne\ is comparable to that of \textsf{FIt-SNE} and much faster
than its Barnes-Hut competitors. Additionally, modern computer processors
generally contain multiple cores, allowing us to make use of multi-threading.
This further reduces the gap between \opentsne\ and \textsf{FIt-SNE}. On the
server-grade processor, \opentsne\ is slightly faster than its pure C++
counterpart when utilizing multiple cores. We speculate that this is could be because \opentsne\ uses
\textsf{numpy} for most linear algebra operations, which may make better use of
the Intel Math Kernel Library (MKL), which is aggressively optimized on
Intel(R) Xeon(R) processors.

Additionally, \opentsne\ provides a flexible API, which allows splitting up the
embedding-construction process into several part, caching slow operations. This
enables users to quickly experiment with different parameter settings and
iterate on their final visualizations.

\subsection*{Ease of Use}

Widespread adoption of novel techniques is almost universally accompanied by
easy-to-use and easy-to-install software. While the t-SNE implementation from
\textsf{scikit-learn} fits this requirement, its implementation is
unfortunately very slow, making it unsuitable for its application to large data
sets. Other C++ implementations such as \textsf{MulticoreTSNE} and
\textsf{FIt-SNE} exhibit better scaling in larger data sets, but do not provide
precompiled binaries, requiring users to compile the software themselves. This
is problematic from a usability standpoint, as many users often lack the
knowledge to do so. This often occurs with Windows operating system, which does
not include a C++ compiler, making the correct configuration cumbersome.

\opentsne\ is a Python package that focuses on accessibility. We provide
precompiled binaries for all major Python versions on all major platforms,
making the installation process as seamless as possible. \opentsne\ can be
installed through the Python Package Index (PyPI) or through conda on the
conda-forge channel. \opentsne\ is designed to be familiar to Python users and
follows the \textsf{scikit-learn} API~\cite{sklearn_api}, which is well
established in the Python data science ecosystem. \opentsne\ implements
multi-threaded versions of both the Barnes-Hut and FIt-SNE approximation
schemes, allowing it to be used on data sets containing millions of data
points. While the Python virtual machine introduces some performance overhead,
the runtime is comparable to that of its C++ counterpart --
\textsf{FIt-SNE}. Finally, \opentsne\ is extensible. Its modular design enables
researchers to quickly experiment with different settings and easily
incorporate their own components into the software. We provide a full feature
list and comparison to other popular t-SNE implementations in
Table~\ref{tab:features}.

\begin{table}
\caption{\label{tab:features}We compare the feature list of \opentsne to three
popular open-source implementations from scikit-learn~(v0.23.1), MulticoreTSNE
(v0.1), and FIt-SNE (v1.1.0). The first section corresponds to packaging and 
distribution. A properly packaged library is easily accessible to users and can
easily be included as a dependency in other software packages. The second
section lists the two approximation schemes. The FIt-SNE approximation scheme
is required for t-SNE to scale up to millions of data points. The final section
provides a list of extensions and improvements to the standard t-SNE algorithm,
many of which can produce markedly better visualizations than under default
parameter settings.}

\newcommand*\rot{\rotatebox{90}}
\renewcommand{\arraystretch}{1.25}

\begin{tabular}{r c c c c c c}
\toprule
\setlength\tabcolsep{6pt}
& \rot{\textsf{scikit-learn}} & \rot{\textsf{MulticoreTSNE}} & \rot{\textsf{FIt-SNE}} & \rot{\textsf{openTSNE}} \\
\toprule
PyPI package & \checkmark & \checkmark & & \checkmark \\
conda package & \checkmark & & & \checkmark \\
Precompiled binaries & \checkmark & & & \checkmark \\
\hline
Barnes-Hut ($\mathcal{O}(N \log N)$) & \checkmark & \checkmark & & \checkmark \\
FIt-SNE ($\mathcal{O}(N)$) & & & \checkmark & \checkmark \\
\hline
Multiscale Gaussian kernels & & & \checkmark & \checkmark \\
Fully-custom affinity kernels & & & & \checkmark \\
Variable degrees of freedom & & & \checkmark & \checkmark \\
Variable exaggeration & & & \checkmark & \checkmark \\
Better initialization & & & \checkmark & \checkmark \\
Automatic learning rate & & & \checkmark & \checkmark \\
Embedding new samples & & & & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\section*{Conclusion}
Text for this section \ldots

\section*{Methods}

This section is meant to familiarize the reader with the t-SNE algorithm and
its recent extensions. We motivate each extension and introduce standard notation.
This brief review is by no means comprehensive, but is
self-contained and introduces the mathematics upon which the \opentsne\ library
is based.

\subsection*{Preliminaries}

t-distributed stochastic neighbor embedding (t-SNE) is a non-linear
dimensionality reduction method, which aims to find a low-dimensional
embedding, where local neighborhoods are preserved. More formally, given a
multi-dimensional data set $\mathbf{X} = \left \{ \mathbf{x}_1, \mathbf{x}_2,
\dots, \mathbf{x}_N \right \} \in \mathbb{R}^D$ where $N$ is the number of data
points in the data set, t-SNE aims to find a low dimensional embedding
$\mathbf{Y} = \left \{ \mathbf{y}_1, \mathbf{y}_2, \dots, \mathbf{y}_N \right
\} \in \mathbb{R}^d$ where $d \ll D$, such that if points $\mathbf{x}_i$ and
$\mathbf{x}_j$ are close in the high-dimensional space, their corresponding
embeddings $\mathbf{y}_i$ and $\mathbf{y}_j$ are also close. Since t-SNE is
primarily used as a visualization tool, $d$ is typically set to two. The
similarity between two data points in the high-dimensional space is defined as

\begin{equation}
p_{j \mid i} = \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / \sigma_i^2 \right )}
{\sum_{k \neq i } \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_k ) / \sigma_i^2 \right )}, \quad p_{i \mid i} = 0
\label{eq:gaussian_kernel}
\end{equation}

\noindent where $\mathcal{D}$ is some distance measure. This is then
symmetrized to

\begin{equation}
p_{ij} = \frac{p_{j \mid i} + p_{i \mid j}}{2N}.
\label{eq:symmetrize}
\end{equation}

The bandwidth of each Gaussian kernel $\sigma_i$ is selected such that the
perplexity of the distribution matches a user-specified parameter value

\begin{equation}
\text{Perplexity} = 2^{H(P_i)}
\end{equation}

\noindent where $H(P_i)$ is the Shannon entropy of $P_i$,

\begin{equation}
H(P_i) = -\sum_i p_{j \mid i} \log_2 (p_{j \mid i}).
\end{equation}

\noindent Different bandwidths $\sigma_i$ enable t-SNE to adapt to the varying
density of the data in the multi-dimensional space. Perplexity is sometimes
interpreted as the continuous analog to the number of nearest neighbors to which
distances will attempt to be preserved.

The similarity between points $\mathbf{y}_i$ and $\mathbf{y}_j$ in the
embedding space is defined using the $t$-distribution with a single degree of
freedom (Cauchy kernel)

\begin{equation}
q_{ij} = \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 \right )^{-1}}
{\sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2 \right )^{-1}},
\quad q_{ii} = 0.
\label{eq:cauchy_kernel}
\end{equation}

The Kullback-Leibler (KL) divergence is used as a measure of agreement
between distributions $\mathbf{P}$ and $\mathbf{Q}$

\begin{equation}
C = \text{KL}(\mathbf{P} \mid \mid \mathbf{Q}) = \sum_{ij} p_{ij} \log \frac{p_{ij}}{q_{ij}}.
\label{eq:kl_divergence}
\end{equation}

\noindent The objective is to find embeddings $\mathbf{Y}$ such that the
KL divergence is minimized.

The corresponding gradient takes the form
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \sum_{j \neq i} \left ( p_{ij} - q_{ij} \right ) \left ( \mathbf{y}_i - \mathbf{y}_j \right ) w_{ij},
\label{eq:tsne_gradient}
\end{equation}
where $w_{ij} = \left ( 1 + || \mathbf{y}_i - \mathbf{y}_j || ^2 \right )^{-1}$
and represents the unnormalized $q_{ij}$.

Optimization is performed with batch gradient descent using the delta-bar-delta
update rule~\cite{jacobs1988increased}. Originally, t-SNE was run for 1000
iterations consisting of two phases: in the first \textit{early exaggeration}
phase, the attractive forces between data points is increased by some factor
$\rho$, typically set to $12$, so that points in the embedding can more easily move throughout the embedding to find
their respective neighbors. The remaining 750 iterations are run with $\rho=1$,
which reverts the attractive forces to their original values and produces the
final embedding.

Belkina \textit{et al.} later found that faster convergence can be achieved by
increasing the learning rate to $\eta=N/12$~\cite{belkina2019automated}. As a side-effect, embeddings converge faster, and the
number of iterations can be lowered to 750, decreasing the overall runtime.
This convention has been adopted by most modern t-SNE implementations.

\subsection*{Efficient Approximation Schemes} A direct evaluation of t-SNE
gradients requires $\mathcal{O}(N^2)$ operations, which makes its application
impractical to any reasonably-sized data set and beckons for the development of
efficient approximation schemes. Van der Maaten observed that the t-SNE
gradient may be cast as an N-body problem where data points represent
particles which attract and repel each
other~\cite{van2014accelerating}. The gradient from Eqn. (\ref{eq:tsne_gradient}) can be rewritten as
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \left [ \sum_{j \neq i} p_{ij} q_{ij} Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) -\sum_{j \neq i} q_{ij}^2 Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) \right ], \label{eq:grad_attr_rep}
\end{equation}
where $Z = \sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2
\right )^{-1}$. This equation can then be viewed as a particule simulation, where 
the two terms represent the attractive and repulsive
forces between individual particles. Each term lends itself to efficient
approximations, which enable us to greatly reduce the time complexity of t-SNE.

\subsubsection*{Attractive Forces}

Van der Maaten observed that evaluating the attractive forces between all pairs
of data points is excessive, and that considering only a handful of nearest
neighbors at each point is sufficient to obtain a good approximation~\cite{van2014accelerating}.
Because distances are transformed to similarities using a Gaussian kernel, and
the bandwidth of each kernel is selected such that only a predefined number of
neighbors fall within the main probability mass of the bell curve through the perplexity parameter, the remaining
data points fall into the exponentially decaying tails of each Gaussian distribution.
These data points are then assigned near-zero probabilities and do not contribute
to the overall attractive forces of data points.
Therefore it is sufficient to calculate the attractive forces for only a
small number of nearest neighbors instead of all $N$ points. By
utilizing tree-based nearest-neighbor search methods, the time complexity is
thus reduced to $\mathcal{O}(N \log N)$. Linderman \textit{et al.} further
realized that, qualitatively, embeddings are visually
indistinguishable when using only \textit{approximate} nearest neighbors,
further reducing time complexity to $\mathcal{O}(N)$~\cite{linderman2019fast}.

\subsubsection*{Repulsive Forces}

Examining the second term of Eqn.~(\ref{eq:grad_attr_rep}) we notice that each
point indiscriminately exerts a repulsive force on all other points. Van der
Maaten proposed an approach based on N-body simulations and used a
space-partitioning Barnes-Hut tree approach to approximate the interaction
between data points~\cite{van2014accelerating}. Briefly, in the 2D case, the approach splits
the space into quadrants and entire regions may be summarized by simple statistics.
If a query point is far away from a given quadrant, the repulsive forces exerted
by all the points in that quadrant onto the query point are summarized by its center of mass. This
reduces the time complexity from $\mathcal{O}(N^2)$ to
$\mathcal{O}(N \log N)$.

More recently, Linderman \textit{et al.} proposed an alternative approach,
FIt-SNE, based on non-uniform convolutions for calculating all pairwise
interactions between repelling data points~\cite{linderman2019fast}. Briefly,
Linderman \textit{et al.} observed that the repulsive forces $\mathbf{R}$ from Eqn. (\ref{eq:grad_attr_rep}) may
be rewritten as \begin{align}
\mathbf{R}_i &= \sum_{j \neq i} q_{ij}^2 Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) \notag \\
&= \sum_{j \neq i} \frac{\mathbf{y}_i - \mathbf{y}_j}{\left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 \right )^{2}}
\bigg/
\sum_{k \neq l} \frac{1}{1 + || \mathbf{y}_k - \mathbf{y}_l || ^2} \label{eq:fitsne_rep}
\end{align} 
and computed by evaluating three terms
\begin{align}
\phi_{1, j} &= \sum_{j \neq i} \frac{1}{1 + || \mathbf{y}_j - \mathbf{y}_i ||^2}, \notag \\
\phi_{2, j} &= \sum_{j \neq i} \frac{\mathbf{y}_j}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 \right)^2}, \notag \\
\phi_{3, j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 \right)^2}. \label{eq:fitsne_terms}
\end{align}
Then the numerator and denominator of Eqn.~(\ref{eq:fitsne_rep}) can be
computed as $\mathbf{y}_i \phi_{1,j} - \phi_{2,j}$ and $Z = \sum_j \phi_{3,j}$,
respectively. These interactions are calculated by interpolating the terms
through a grid of equispaced interpolation points. This shifts the
computational burden onto the interpolation points and reduces the time
complexity to $\mathcal{O}(N)$.

% -------- pointless...
%\opentsne\ provides efficient implementations for all the aforementioned
%improvements but defaults to using approximate nearest neighbor search and the
%non-uniform convolutions approach as these enable the user to quickly create
%informative visualizations for data sets containing up to millions of samples.

\subsection*{Embedding New Samples}

t-SNE is non-parametric and does not define an explicit mapping from the
high-dimensional space to the embedding space. Therefore embeddings of new data
points need to be found through the use of optimization
techniques~\cite{policar2019embedding}. When adding new data points to an
existing, reference embedding, the reference data points are fixed in place while
new data points are allowed to find their respective positions. The
optimization remains the same as in standard t-SNE with only slight
modifications to $p_{ij}$ and $q_{ij}$
\begin{align}
p_{j \mid i} &= \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) /  \sigma_i^2 \right )}{\sum_{i} \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) / \sigma_i^2 \right )}, \\
q_{j \mid i} &= \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}}{\sum_{i}\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}},
\end{align}
\noindent where $\mathbf{V} = \left \{ \mathbf{v}_1, \mathbf{v}_2, \dots,
\mathbf{v}_M \right \} \in \mathbb{R}^D$ where $M$ is the number of samples in
the new data set and $\mathbf{W} = \left \{ \mathbf{w}_1, \mathbf{w}_2, \dots,
\mathbf{w}_M \right \} \in \mathbb{R}^d$. Additionally, we omit the
symmetrization step in Eqn.~(\ref{eq:symmetrize}).
Plugging these
The gradients of
$\mathbf{w}_j$ with respect to the loss (Eqn.~\ref{eq:kl_divergence}) are:
\begin{equation}
\frac{\partial C}{\partial \mathbf{w}_j} = 2 \sum_i \left ( p_{j \mid i} - q_{j \mid i} \right ) \left ( \mathbf{y}_i - \mathbf{w}_j \right ) \left ( 1 + || \mathbf{y}_i - \mathbf{w}_j || ^2 \right )^{-1}.
\label{eq:gradient}
\end{equation}

As in standard t-SNE, a direct calculation of gradients takes $\mathcal{O}(N
\cdot M)$ time, but it is straightforward to adapt the Barnes-Hut and FIt-SNE
approximation schemes, reducing the time complexity to $\mathcal{O}(M \log N)$
and $\mathcal{O}(\max \{ N, M \})$, respectively. In the FIt-SNE approximation
scheme, we additionally exploit the fact that the reference embedding remains
fixed throughout the optimization of newly added points, and precompute the
interpolation grid. This further reduces the runtime complexity from
$\mathcal{O}(\max \{ N, M \})$ to $\mathcal{O}(M)$.

\subsection*{Alternative Perplexity Kernels}

In standard t-SNE, distances are converted to similarities through the use of
Gaussian kernels of varying bandwidths. The bandwidths are indirectly
determined by the user-specified perplexity parameter so that a fixed number of
nearest data points will have non-zero similarity. One common trick for
uncovering the global relations between clusters is to increase perplexity so
that more long-range interactions are preserved in the final embedding.
However, one unfortunate side effect of increasing perplexity is that smaller
clusters get absorbed into larger ones.

Kobak \& Berens suggest that replacing the Gaussian kernel
with a mixture of Gaussians may provide better insight into both the local and
global structure~\cite{kobak2019art}. Therefore, the similarities between data points in the input
space are computed with

\begin{align}
  p_{j\mid i} &\propto \frac{1}{\sigma_{1,i}} \exp \left ( -\mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / 2\sigma_{1,i}^2 \right ) \notag \\
  &+ \frac{1}{\sigma_{2,i}} \exp \left ( -\mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / 2\sigma_{2,i}^2 \right ).
\end{align}

The bandwidth of each Gaussian $\sigma_{1,i}$ and $\sigma_{2,i}$ is determined
by selecting different perplexity values. Using a kernel with
perplexity 500 captures long-range interactions which preserve global
cluster organization. Combining these with a more narrow kernel with
perplexity 50 prevents small, well-defined clusters from being absorbed
into larger ones, leading a better overall snapshot into the data structure.

\subsection*{Variable Degrees of Freedom}

Standard t-SNE reveals the clustering structure at a single level of
resolution. Different perplexity parameter values can be used to identify
global cluster relationships or small, well-isolated groups. Unfortunately,
varying perplexity values can be time-consuming as this involves recomputing
the KNN graph which is often the most expensive part of the t-SNE algorithm.
Alternatively, Kobak \textit{et al.}~\cite{kobak2019heavy} suggest that varying
the degree of freedom in the t-distribution can be used to explore the clustering structure at different
levels of resolution.

Standard t-SNE models similarities between data points in the embedding space
using a t-distribution with a single degree of freedom

\begin{equation}
q_{ij} \propto \left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 / \alpha \right )^{-\alpha} = \frac{1}{\left( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 / \alpha \right)^\alpha }.
\end{equation}
In standard t-SNE $\alpha=1$ so this simplifies to the Cauchy kernel from
Eqn.~(\ref{eq:cauchy_kernel}). The gradient of the loss function then becomes
\begin{align}
\frac{\partial C}{\partial \mathbf{y}_i} &= 4 \sum_{j \neq i} \left ( p_{ij} - q_{ij} \right ) w_{ij}^{1/\alpha} \left ( \mathbf{y}_i - \mathbf{y}_j \right ),
\end{align}
which can, again, be cast as attractive and repulsive forces between particles
\begin{align}
\frac{\partial C}{\partial \mathbf{y}_i} &= 4 \sum_{j \neq i} p_{ij} w_{ij}^{1/\alpha} (\mathbf{y}_i - \mathbf{y}_j) \notag \\
&- 4 \sum_{j \neq i} w_{ij}^{\frac{\alpha+1}{\alpha}} / Z (\mathbf{y}_i - \mathbf{y}_j).
\end{align}

Adapting existing approximation schemes to this formulation is straightforward.
\opentsne\ provides efficient implementations of both the Barnes-Hut and the FIt-SNE
approximation schemes, where we modify the terms from
Eqn.~(\ref{eq:fitsne_terms}) to

\begin{align}
\phi_{1,j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)^{\alpha+1}}, \notag \\
\phi_{2,j} &= \sum_{j \neq i} \frac{\mathbf{y}_j}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)^{\alpha+1}}, \notag \\
\phi_{3,j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)}. \notag
\end{align}

\subsection*{Variable Exaggeration}

Embeddings produced by standard t-SNE often produce clusters separated by thin
boundaries and make use of all available space. While this is desirable at times,
it often obscures the global relationships between clusters as all
neighboring clusters appear about the same distance from one another. Other
dimensionality reduction methods such as UMAP~\cite{2018arXivUMAP} or
ForceAtlas2~\cite{jacomy2014forceatlas2} tend to produce embeddings where
clusters appear more compact and the white-space separating the clusters may be
interpreted at least partially as a loose measure of distance.

B{\"o}hm \textit{et al.} showed that the exaggeration factor
$\rho$ can be used to produce layouts more similar to UMAP and ForceAtlas2~\cite{bohm2020unifying}.
By incorporating exaggeration into later phases of optimization, t-SNE is able
introduce more white-space between clusters and better reflect the global
relations between clusters. B{\"o}hm \textit{et al.} found that using $\rho=4$ and $\rho=30$
produces embeddings visually similar to UMAP and ForceAtlas2, respectively.

While it is difficult to claim one is better than the other, different settings
of $\rho$ may uncover different properties and clustering structures. For
example, in single-cell data, standard t-SNE with $\rho=1$ often uncovers
distinct groups of cell-types but obscures developmental paths. This problem is
exacerbated when dealing with large numbers of data points. On the other hand,
ForceAtlas2 has often been used to uncover trajectories and transitions between
cell types, but larger clusters often absorb small, distinct groups of cells.
Therefore one may run t-SNE with two different exaggeration settings, one to
highlight the developmental transitions between cell types and one to identify
clear populations of cells.

\subsection*{Globally Consistent Initialization Schemes}

t-SNE, UMAP, and ForceAtlas2 can all be cast as force-directed layout algorithms,
which operate on the $k$-nearest neighbor graph. Each method constructs the graph
and specifies the attractive and repulsive forces in its own way, but ultimately,
point positions are found by balancing the attractive and repulsive forces between data points.

The final positioning of data points using this class of algorithms is largely
dependent on the embedding initialization. UMAP found early success as it was
able to produce visualizations that better captured the global organization of
clusters. This was due to its initialization scheme, which initialized point
positions using Laplacian eigenmaps. On the other hand, most implementations of
t-SNE performed random initialization, which resulted in poor global coherence.
However, Kobak \& Linderman recently showed that using the same initialization
for both methods leads to visualizations, which exhibit similarly good global
coherency. Conversely, when initialized randomly, both methods produce
visualizations where clusters are arbitrarily positioned in the embedding space.

\opentsne\ defaults to using the two leading principal components as
initialization, but also provides a spectral approach similar to UMAP.
This allows clusters in resulting embeddings to be organized in a more globally coherent
manner, leading to increased visualization interpretability.

\section*{Availability}

openTSNE is distributed under the BSD-3-Clause License and is publicly
available as an open-source package at
\url{https://github.com/pavlin-policar/openTSNE}. openTSNE is also available on
PyPI and conda-forge. The data sets used and scripts used in this study are
included in accompanying notebooks, publicly available at
\url{https://github.com/pavlin-policar/opentsne-paper}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
We would like to thank Dmitry Kobak for various helpful discussions and best
	practices when using t-SNE, as well as his contributions to the source
	code. We would also like to thank George Linderman for his help with
	the FIt-SNE algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{references}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

\end{backmatter}
\end{document}