### 2020-10-01 paper/policar2020-opentsne.tex
revision of the main text

\documentclass[twocolumn]{bmcart}
%\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\usepackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

\usepackage{graphicx}
\usepackage{minted}
\usemintedstyle{vs}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\startlocaldefs
\newcommand{\opentsne}{\textsf{openTSNE}}
\endlocaldefs

\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Software}

\title{openTSNE: a modular Python library for t-SNE dimensionality reduction and embedding}

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={pavlin.policar@fri.uni-lj.si}   % email address
]{\inits{PGP}\fnm{Pavlin G.} \snm{Poli\v{c}ar}}
\author[
   addressref={aff1,aff2}
]{\inits{MS}\fnm{Martin} \snm{Stra\v{z}ar}}
\author[
   addressref={aff1,aff3}
]{\inits{BZ}\fnm{Bla\v{z}} \snm{Zupan}}

\address[id=aff1]{%
  \orgname{Faculty of Computer and Information Science, University of Ljubljana},
  %\street{Ve\v{c}na Pot 113},
  \postcode{SI 1000}
  \city{Ljubljana},
  \cny{Slovenia}
}
\address[id=aff2]{%
  \orgname{Broad Institute of Harvard and MIT},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{MA 02142}
  \city{Cambridge},
  \cny{U.S.A.}
}
\address[id=aff3]{%
  \orgname{Department of Molecular and Human Genetics, Baylor College of Medicine},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{TX 77030}
  \city{Houston},
  \cny{U.S.A.}
}

\end{fmbox}% comment this for two column layout

\begin{abstractbox}
\begin{abstract}
Point-based visualizations of large, multi-dimensional data from molecular
biology can reveal meaningful clusters. One of the most popular
techniques to construct such visualizations is $t$-distributed stochastic
neighbor embedding (t-SNE). Several extensions of t-SNE have
recently been proposed to address issues of scalability and the quality
of the resulting visualizations. We introduce \opentsne, a modular
Python library that implements the core t-SNE algorithm and its
extensions. The library is orders of magnitude faster than existing
popular implementations, including those from scikit-learn. Unique to
\opentsne\ is also mapping new data to existing embeddings,
which can surprisingly assist in solving batch effects.
\end{abstract}

\begin{keyword}
\kwd{t-SNE}
\kwd{embedding}
\kwd{visualization}
\kwd{dimensionality reduction}
\kwd{Python library}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

\section*{Background}

The abundance of high-dimensional data sets in molecular biology calls for dimensionality reduction techniques, particularly for methods that produce informative data visualizations. Popular approaches include 
principal component analysis (PCA), multidimensional scaling, t-distributed stochastic neighbor embedding (t-SNE)~\cite{maaten2008visualizing}, and uniform manifold approximation and projections (UMAP)~\cite{2018arXivUMAP}. Among these, t-SNE lately received much attention as it can address high volumes of data and reveal a meaningful clustering structure. Increased
resolution and throughput of modern molecular assays has lead to the frequent use of t-SNE in diverse fields including, but not limited to, single-cell transcriptomics(scRNA-seq,~\cite{macosko2015highly,cao2019single,tasic2018shared}),
human genetics~\cite{hirata2019genetic}, metagenomic
assembly~\cite{beaulaurier2018metagenomic}, the spatial organization of microbial communities~\cite{sheth2019spatial} and metabolomics~\cite{tkachev2019differences}. Reports on single-cell gene expression data, our running example, often start with an overview of the cell landscape, where t-SNE embeds high-dimensional expression profiles into a two-dimensional
space. Figs.~\ref{fig:macosko}.a and \ref{fig:macosko}.b show two such embeddings.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{macosko2015}
\caption{\label{fig:macosko}
Three t-SNE embeddings demonstrate theoretical advances and corresponding implementations in
\opentsne. The data in \textbf{(a)} and \textbf{(b)} represent 44,808
single-cell gene-expression profiles of mouse retinal cells from Macosko \textit{et al.}~\cite{macosko2015highly}. The data in \textbf{(c)} contains 27,499 expression profiles from mouse
retinal cells from Shekhar \textit{et
al.}~\cite{shekhar2016comprehensive}. \textbf{(a)} We construct a t-SNE
embedding following the parameter choices from the original publication
by Maaten \& Hinton~\cite{maaten2008visualizing}. The visualization
shows no preservation of the global organization of clusters,
resulting from random initialization and an affinity model focused on
preserving local neighborhoods. \textbf{(b)} A modern t-SNE
embedding, utilizing the latest theoretical advances and practical
recommendations constructed using a multi-scale
affinity model, preserving both short-range and long-range interactions
between data points and initialized so that the global layout is
meaningful. Unlike in \textbf{(a)}, the green and blue clusters
representing different sub-types of amacrine and bipolar cells are now
localized to the same regions of the space, indicating a higher level
of similarity than to other cell types. Embedding in \textbf{(c)} shows how existing
t-SNE reference atlases can be used to place new samples into existing
embeddings. The positions of new data points correspond to cell types
from the reference atlas.
}
\end{figure*}

Despite its utility, t-SNE has often been criticized for its limited scalability, lack of global organization -- t-SNE emphasizes
local clusters that are arbitrarily scattered in the low-dimensional space --
and the absence of theoretically-founded methods to map new data into
existing embeddings~\cite{ding2018interpretable,becht2019dimensionality}. Most
of these shortcomings have recently been addressed. Linderman \textit{et al.}
sped-up the method through an interpolation-based approximation, achieving linear complexity in the number of
samples~\cite{linderman2019fast}. Kobak \& Berens proposed several techniques
to improve global positioning, including estimating similarities with a mixture
of Gaussian kernels~\cite{kobak2019art}. In our previous work, we presented a principled approach for embedding new samples into existing
visualizations~\cite{policar2019embedding}.

\section*{Results}

We introduce \opentsne, a comprehensive Python library that implements t-SNE
and all its recently proposed extensions, including:
\begin{enumerate}
\item the implementation of efficient approximation 
	schemes~\cite{van2014accelerating,linderman2019fast} allowing the embedding
	of millions of data points,
\item the addition of new data samples into a fixed existing
	embeddings~\cite{policar2019embedding},
\item improved initialization schemes~\cite{kobak2019umap} leading to more
	globally consistent layouts,
\item variable degrees of freedom~\cite{kobak2019heavy} allowing the inspection
	of data at different levels of resolution,
\item multi-scale similarity kernels~\cite{kobak2019art} which preserve small,
	well-defined clusters and uncover global relationships between clusters,
	and
\item improved defaults for learning rate and number of
	iterations~\cite{belkina2019automated} which produce embeddings at a lower
	computational cost.
\end{enumerate}

The \opentsne\ is compatible with the Python data science ecosystem and libraries that include
\textsf{numpy}, \textsf{scikit-learn}, \textsf{scanpy}), thus providing a familiar and intuitive application program interface to the new users. Its modular design encourages extensibility and experimentation with various settings and changes in the analysis pipeline.
We have released \opentsne\ in an open-source to support adaptations by the wider community. At the time of the writing, the library averages several thousands of weekly downloads and has received almost 700 GitHub stars, an appreciation score gained by only $0.3\%$ of public Python repositories. The source code is available at
\url{https://github.com/pavlin-policar/openTSNE} and supports installations from
\textsf{PyPI} and \textsf{conda-forge}, the two most widely adopted Python
package managers.

Accessibility of the latest theoretical advances in t-SNE is one of the core design principles of \opentsne. These advances are accessible through a programming interface of the library, where we closely follow the style of the interface as implemented in
\textsf{scikit-learn}~\cite{sklearn_api}, a popular library for machine learning. The following snippet exemplifies this interface and codes for four visualizations in Fig.~\ref{fig:tasic}, demonstrating the effect of a particular recent theoretical advance in the t-SNE algorithm.

\begin{minted}[
fontsize=\footnotesize,
frame=lines,
framesep=2mm,
baselinestretch=1.1
]{python}
import openTSNE
from openTSNE.affinity import Multiscale
# a - standard t-SNE
tsne1 = openTSNE.TSNE(perplexity=30).fit(X)
# b - high perplexity  for globa structure
tsne2 = openTSNE.TSNE(perplexity=500).fit(X)
# c - multiscale kernel for local/global structure
aff = Multiscale(X, perplexities=[30, 500])
tsne3 = openTSNE.TSNE(affinities=aff).fit(X)
# d - decrease dof for higher resolution
tsne4 = openTSNE.TSNE(dof=0.6).fit(X)
\end{minted}

\noindent The code snippet considers a \textsf{numpy} array or
\textsf{scipy} sparse matrix \texttt{X} and creates four
\texttt{TSNEEmbedding} objects. These objects encode t-SNE embeddings that may subsequently be optimized under different parameter settings or used to embed new samples into the embedding landscape. Function calls may include additional parameters not shown in the example code but explained in accompanying notebooks available at \url{https://github.com/pavlin-policar/opentsne-paper}.

\section*{Discussion}

\subsection*{Uncovering Structure in High-Dimensional Data}

Dimensionality reduction techniques implicitly assume that high-dimensional data lies on a lower-dimensional manifold, which can accurately be captured by a smaller number of dimensions. However, there is no evidence that every data set can accurately be described by using only two dimensions, and any such embedding will inevitably lead to a loss of information. Thus, it is beneficial to examine multiple embeddings, each of which provides a different perspective on topology and other data characteristics.

\begin{figure*}[htbp]
	\includegraphics[width=\textwidth]{tasic2018}
	\caption{\label{fig:tasic}We use openTSNE to create four different
    visualizations of the Tasic \textit{et al.}~\cite{tasic2018shared} data,
    each providing a different perspective into the topology of the data.
    The data set contains 21,874 single-cells originating from the mouse
	neocortex. Cluster annotations and colors are taken from the original
	publication. Warm colors correspond to excitatory neurons, cool colors
	correspond to inhibitory neurons, and gray/brown colors correspond to
	non-neuronal cells. The standard t-SNE \textbf{(a)} emphasizes local
	structure while increasing perplexity \textbf{(b)} results in a more
	meaningful layout of the clusters. We can also combine the two
	perplexities by using a multiscale kernel affinity model \textbf{(c)}
	and obtain a good trade-off between global and local structure.
	Alternatively, we can inspect more fine-grained structure and reveal
	smaller clusters by using a more heavy-tailed kernel \textbf{(d)}.}
\end{figure*}

We illustrate this point by generating four different embeddings of the data on single-cell gene expression in mouse brain~\cite{tasic2018shared}.  Fig.~\ref{fig:tasic}.a shows an embedding using default t-SNE parameters. While different clusters of excitatory and inhibitory neurons appear close to one another, all clusters appear equidistant from their neighbors, and the overall relations between groups are not obvious. Embedding in Fig.~\ref{fig:tasic}.b focuses on preserving larger neighborhoods of points, resulting in a more globally consistent layout where relations between clusters become more apparent. Here, it is evident from the increased white space between groups that there is one large class of excitatory neurons and two related classes of inhibitory neurons. Unfortunately, focusing on preserving large neighborhoods leads to the absorption of smaller clusters by the larger ones. Alternatively, Fig.~\ref{fig:tasic}.c uses multi-scale similarity kernels that aim to preserve both the global organization of clusters and prevent the cluster absorption. We constructed the embeddings from Fig.~\ref{fig:tasic}.d with the settings used for Fig.~\ref{fig:tasic}.a, but at a finer level of resolution. The distinction between these two depictions reveals that many clusters are composed of subgroups representing different cell subtypes.

When dealing with data containing millions of data points, standard t-SNE embeddings often become unwieldy -- cluster boundaries are blurred, large clusters absorb smaller ones, and relationships between clusters become increasingly difficult to interpret. We constructed Fig.~\ref{fig:cao}.a from the data containing expression profiles of over two million single cells captured at different time points in mouse development. The embedding indicates numerous clusters with transitions between time points, as marked by the color-coding, that are difficult to interpret. Kobak \& Berens observed that increasing attractive forces between similar data points controlled via the \textit{exaggeration} parameter leads to more compact clusters, and subsequently, more informative visualizations~\cite{kobak2019art}. For instance, \ref{fig:cao}.b doubles the default exaggeration, which uncovers some of the data's overall structure. Further doubling the exaggeration in \ref{fig:cao}.c allows us to observe that the data is comprised of two main groups of cells and eight somewhat smaller clusters. The visualization also reveals several tiny clusters, possibly corresponding to
rare cell types.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{cao2019}
  \caption{\label{fig:cao}
Increasing the exaggeration parameter leads to compact clusters, highlighting the data's global organization and emphasizing the continuity between cell states. The data set from Cao \textit{et al.}~\cite{cao2019single} contains expression profiles from 2,058,652 single cells. The data were collected from mice embryos at different development stages at daily intervals after 9.5 to 13.5 days. \textbf{(a)} reveals that the data is comprised of two main components -- the neural tube and mesenchymal cells -- as well as several other smaller clusters. Additionally, the colors indicate developmental progression with red indicating least-developed cells and blue indicating most developed cells. The overall developmental trajectory is most apparent with higher exaggeration levels, showing red cells slowly transitioning into blue cells. Progressively easing the exaggeration factor uncovers finer clusters within the larger groups, as shown in
\textbf{(b)} with exaggeration of two and subsequently in \textbf{(c)}, where we show the standard t-SNE with no exaggeration. 32,011 putative doublets are excluded from the visualizations.
}
\end{figure*}

Exaggeration can highlight transitions between cell states in developmental studies. Standard t-SNE often produces embeddings with clearly defined, discrete clusters. We can adjust the level of granularity and resolution of the clusters with several parameters in Fig.~\ref{fig:tasic}. However, discrete clusters are often undesired in developmental studies where cells' state is changed continuously. To this end, other embedding techniques such as UMAP and ForceAtlas2~\cite{jacomy2014forceatlas2} tend to capture the continuity between cell states better. Recently, Kobak \textit{et al.}~\cite{TODO} have shown that embeddings produced by t-SNE with exaggeration values of 4 and $\sim30$ construct embeddings which are markedly similar to UMAP and ForceAtlas2, respectively. For example, in Fig.~\ref{fig:cao}.a, the developmental trajectory between different time points is difficult to observe due to many sprawled out clusters. On the other hand, it is easier to trace the development when we increase the exaggeration factor from $1$ to $4$ (Fig.~\ref{fig:cao}.b-c).

\subsection*{Embedding New Samples}

Unlike other popular dimensionality reduction methods such as principal
component analysis or autoencoders, t-SNE is a non-parametric method and does not define an explicit mapping to the embedding space. Therefore embeddings of new data points need to be found through optimization~\cite{policar2019embedding}. \opentsne\ is currently the only publicly available library allowing users to add new samples to existing embeddings.

Figs.~\ref{fig:macosko}.b and \ref{fig:macosko}.c demonstrate how we can use a previously labeled single-cell data set and embed cells from a separate experiment into the reference landscape. The reference data from Macosko~\textit{et al.}~\cite{macosko2015highly} contains gene expression profiles from mouse retinal cells. By embedding the samples from a similar experiment on bipolar retinal cells by Shekhar~\textit{et al.}~\cite{shekhar2016comprehensive}, we can correctly map the bipolar cell clusters in the reference.

Embedding single cells into existing reference atlases can also be useful for cell-type classification in cases of unknown cell identities. For instance, in Fig.~\ref{fig:transform}, we construct a reference embedding using labeled data from Hochgerner~\textit{et al.}~\cite{hochgerner2018conserved} containing gene-expression profiles of cells from the mouse brain. The authors assign a type to each cell. We can verify their classification accuracy by visualizing the expression of well-established gene markers for the major cell types. We then embed cells from Harris~\textit{et al.}~\cite{harris2018classes} into a constructed atlas. In Harris~\textit{et al.}~, labels are provided only for neuronal cells. In the resulting mapping, we can quickly identify other non-neuronal cell types, including oligodendrocytes and astrocytes. We can further use marker genes to validate that the mapping in the reference landscape was correct.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{transform_hochgerner}
  \caption{\label{fig:transform}
\openTSNE supports embedding new samples into an existing reference t-SNE landscape. For the series of visualizations shown in this figure, we first construct a t-SNE embedding for the data from Hochgerner \textit{et al.}~\cite{hochgerner2018conserved} containing 24,185 developing, single cells from the mouse hippocampus. The data contains gene expression in different neurons, supporting glia, and other vascular cells (upper left). Data points representing cells are colored according to cell-types assigned in the original publication; see the legend from Fig.~\ref{fig:tasic} to map colors to cell-type. We then embed new, hippocampal cells collected in a study by Harris \textit{et al.}~\cite{harris2018classes} using the embedding of Hochgerner \textit{et al.} data as a reference. In their study, Harris \textit{et al.} collected 6,971 single-cells and focused on identifying different types of inhibitory neurons. However, almost half of the collected cells are not neurons and were left uncharacterized. Inspecting the embeddings of these cells in the reference embedding (bottom left) reveals that in addition to inhibitory neurons, the data contains several supporting glial cells as well as a small population of endothelial cells. We can verify our approach's accuracy by inspecting marker genes for the major cell types in the reference (top row) and embedded samples (bottom row).
}
\end{figure*}

The examples presented above demonstrate how to use \opentsne\ to quickly gain insight into a newly-sequenced, single-cell data sets by utilizing existing cell atlases. The approach is general and not limited to single-cell gene expression, and one can, in principle, apply it to any tabular data set regardless of field.

\subsection*{Versatility}

Versatility, the ability to use and combine different optimization approaches to construct the embedding spaces, is another of \opentsne's core design principles. Kobak \& Berens recently introduced several recommendations and tricks to obtain better and more meaningful t-SNE visualizations~\cite{kobak2019art}. These include multi-scale similarity kernels, perplexity annealing, and increasing exaggeration when working with more massive data sets. \opentsne\ provides a flexible program interface to incorporate these improvements in just a few code lines. Furthermore, \opentsne\ supports custom affinity models, enabling users to construct t-SNE embeddings on non-tabular relational data: the only requirement imposed by the affinity-model is the availability of similarity between data points, that is, the distance matrix.  Finally, the \opentsne's comprehensive callback system can be utilized to monitor and adapt different stages of the optimization phase and has been used to construct visually appealing animations of the t-SNE optimization process.

\subsection*{Speed}

One of the t-SNE's common criticisms is limited scalability to large data sets containing, for instance, millions of data points~\cite{becht2019dimensionality}. The culprit for slow response time stems from a particular optimization procedure and its specific implementation in the popular Python library. Namely, until quite recently, most popular implementations of t-SNE were based on the Barnes-Hut approximation scheme developed by van der Maaten in 2014~\cite{van2014accelerating} with asymptotic complexity $\mathcal{O}(N \log N)$, where $N$ is the number of data items ({\em e.g.} cells). The most widely-used implementation of t-SNE came from \textsf{scikit-learn}, which exhibits long runtimes when compared to its C++ counterpart -- \textsf{MulticoreTSNE}. The multi-threaded \textsf{MulticoreTSNE} implementation~\cite{Ulyanov2016} can construct t-SNE embeddings of millions of data points in a matter of hours on widely-accessible, consumer-grade processors (Fig.~\ref{fig:benchmarks}). However, recently, Linderman \textit{et al.} developed a new approximation scheme -- FIt-SNE -- which further reduces the asymptotic time complexity to $\mathcal{O}(N)$. We have included this scheme in \opentsne\, enabling the embedding of large data sets in a matter of minutes.

Fig.~\ref{fig:benchmarks} benchmarks four Python t-SNE implementations, including those from \textsf{scikit-learn} (v0.23.1), \textsf{MulticoreTSNE} (v0.1), \textsf{FIt-SNE} (v1.1.0), and our \opentsne\ (v0.4.3). We perform benchmarks on two computational platforms, one representing the usage on a personal computer and the other utility of these libraries on a high-performance computing platform. The Intel(R) Core i7 is commonly found in consumer-grade laptop computers, while Intel(R) Xeon(R) processors appear in high-performance computing machines. Benchmarks were run for $1,000$ iterations with the original t-SNE parameters, as some implementations do not allow their modification.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{benchmarks}
  \caption{\label{fig:benchmarks}
We benchmark openTSNE (v0.4.3) against three popular open-source implementations from scikit-learn~\cite{pedregosa2011scikit} (v0.23.1), MulticoreTSNE~\cite{Ulyanov2016} (v0.1), and FIt-SNE~\cite{linderman2019fast} (v1.1.0). Experiments were run on a consumer-grade Intel Core i7-7700HQ processor found in laptop computers, and on a server-grade Intel Xeon E5-2650. To generate benchmark data sets of different sizes, we subsampled data from the 10X Genomics 1.3 million mouse brain data set five times, resulting in five different data sets for each size. In total, we run each implementation on 30 different data sets. Notice that openTSNE scales similarly to FIt-SNE, as they both use the same interpolation-based approximation scheme, while scikit-learn and MulticoreTSNE utilize the Barnes-Hut approximation.
}
\end{figure*}

Benchmark results (Fig.~\ref{fig:benchmarks}) confirm that both \textsf{FIt-SNE} and \opentsne\ scale better than their Barnes-Hut counterparts -- \textsf{scikit-learn} and \textsf{MulticoreTSNE}. While \opentsne\'s implementation uses Python that incurs some runtime overhead compared to its C++ counterpart \textsf{FIt-SNE}, the two libraries' speeds are surprisingly comparable. Modern computer processors contain multiple cores, allowing us to use multi-threading, which further reduces the gap between \opentsne\ and \textsf{FIt-SNE}. On the server-grade processor, \opentsne\ is slightly faster than its pure C++ counterpart when utilizing multiple cores. \opentsne\ uses \textsf{numpy} for most linear algebra operations, which may make better use of the Intel Math Kernel Library (MKL), which is aggressively optimized on Intel(R) Xeon(R) processors.

\opentsne\ provides a flexible API, allowing splitting up the embedding-construction process into several parts, caching slow operations. Running t-SNE optimization in stages enables users to quickly experiment with different parameter settings and iterate on their final visualizations.

\subsection*{Ease of Use}

Intuitive access and simple installation procedures almost universally correlate with the widespread adoption of novel computational techniques. While the t-SNE from \textsf{scikit-learn} fits this requirement, its implementation is prohibitively slow for even moderately-sized data sets that span tens of thousands of data records. Other C++ implementations such as \textsf{MulticoreTSNE} and \textsf{FIt-SNE} exhibit better scaling in more massive data sets, but do not provide precompiled binaries and require users to compile the software themselves. This problem is critical, for instance, for users of Windows, where the C++ compiler does not come with the system, making the correct configuration of current t-SNE implementations cumbersome.

We have designed \opentsne\ to make it accessible to a broader audience. We provide precompiled binaries for all major Python versions on all major platforms, making the installation process as seamless as possible. One can install \opentsne\ through the Python Package Index (PyPI) or conda on the conda-forge channel. \opentsne\ interface follows that of \textsf{scikit-learn}, which is well established in the Python data science ecosystem. \opentsne\ implements multi-threaded versions of both the Barnes-Hut and FIt-SNE approximation schemes, enabling it for data sets containing millions of data points. While the Python virtual machine introduces some performance overhead, the runtime is comparable to its C++ counterpart -- \textsf{FIt-SNE}. Finally, \opentsne\ is extensible. Its modular design enables researchers to quickly experiment with different settings and easily incorporate their components into the software. We provide a full feature list and comparison to other popular t-SNE implementations in Table~\ref{tab:features}.

\begin{table}
\caption{\label{tab:features}
We compare features of \opentsne to three popular open-source implementations from scikit-learn~(v0.23.1), MulticoreTSNE (v0.1), and FIt-SNE (v1.1.0). The first section of the comparison addresses packaging and distribution. A properly packaged library is easily accessible to users, and developers should easily include it in dependency lists of other software packages. The second section of the comparison lists the two existing t-SNE approximation schemes. The FIt-SNE approximation scheme is required for t-SNE to scale up to millions of data points. The final section provides a list of extensions and improvements to the standard t-SNE algorithm, many of which can produce markedly better visualizations.
}

\newcommand*\rot{\rotatebox{90}}
\renewcommand{\arraystretch}{1.25}

\begin{tabular}{r c c c c c c}
\toprule
\setlength\tabcolsep{6pt}
& \rot{\textsf{scikit-learn}} & \rot{\textsf{MulticoreTSNE}} & \rot{\textsf{FIt-SNE}} & \rot{\textsf{openTSNE}} \\
\toprule
PyPI package & \checkmark & \checkmark & & \checkmark \\
conda package & \checkmark & & & \checkmark \\
Precompiled binaries & \checkmark & & & \checkmark \\
\hline
Barnes-Hut ($\mathcal{O}(N \log N)$) & \checkmark & \checkmark & & \checkmark \\
FIt-SNE ($\mathcal{O}(N)$) & & & \checkmark & \checkmark \\
\hline
Multiscale Gaussian kernels & & & \checkmark & \checkmark \\
Fully-custom affinity kernels & & & & \checkmark \\
Variable degrees of freedom & & & \checkmark & \checkmark \\
Variable exaggeration & & & \checkmark & \checkmark \\
Better initialization & & & \checkmark & \checkmark \\
Automatic learning rate & & & \checkmark & \checkmark \\
Embedding new samples & & & & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\section*{Conclusion}
Text for this section \ldots

\section*{Methods}

This section is meant to familiarize the reader with the t-SNE algorithm and
its recent extensions. We motivate each extension and introduce standard notation.
This brief review is by no means comprehensive, but is
self-contained and introduces the mathematics upon which the \opentsne\ library
is based.

\subsection*{Preliminaries}

t-distributed stochastic neighbor embedding (t-SNE) is a non-linear
dimensionality reduction method, which aims to find a low-dimensional
embedding, where local neighborhoods are preserved. More formally, given a
multi-dimensional data set $\mathbf{X} = \left \{ \mathbf{x}_1, \mathbf{x}_2,
\dots, \mathbf{x}_N \right \} \in \mathbb{R}^D$ where $N$ is the number of data
points in the data set, t-SNE aims to find a low dimensional embedding
$\mathbf{Y} = \left \{ \mathbf{y}_1, \mathbf{y}_2, \dots, \mathbf{y}_N \right
\} \in \mathbb{R}^d$ where $d \ll D$, such that if points $\mathbf{x}_i$ and
$\mathbf{x}_j$ are close in the high-dimensional space, their corresponding
embeddings $\mathbf{y}_i$ and $\mathbf{y}_j$ are also close. Since t-SNE is
primarily used as a visualization tool, $d$ is typically set to two. The
similarity between two data points in the high-dimensional space is defined as:

\begin{equation}
p_{j \mid i} = \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / \sigma_i^2 \right )}
{\sum_{k \neq i } \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_k ) / \sigma_i^2 \right )}, \quad p_{i \mid i} = 0
\label{eq:gaussian_kernel}
\end{equation}

\noindent where $\mathcal{D}$ is some distance measure. This is then
symmetrized to

\begin{equation}
p_{ij} = \frac{p_{j \mid i} + p_{i \mid j}}{2N}.
\label{eq:symmetrize}
\end{equation}

The bandwidth of each Gaussian kernel $\sigma_i$ is selected such that the
perplexity of the distribution matches a user-specified parameter value

\begin{equation}
\text{Perplexity} = 2^{H(P_i)}
\end{equation}

\noindent where $H(P_i)$ is the Shannon entropy of $P_i$,

\begin{equation}
H(P_i) = -\sum_i p_{j \mid i} \log_2 (p_{j \mid i}).
\end{equation}

\noindent Different bandwidths $\sigma_i$ enable t-SNE to adapt to the varying
density of the data in the multi-dimensional space. Perplexity is sometimes
considered the continuous analog to the $k$ nearest neighbors to which
distances will be preserved.

The similarity between points $\mathbf{y}_i$ and $\mathbf{y}_j$ in the
embedding space is defined using the $t$-distribution with a single degree of
freedom. This is also known as the Cauchy kernel:

\begin{equation}
q_{ij} = \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 \right )^{-1}}
{\sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2 \right )^{-1}},
\quad q_{ii} = 0.
\label{eq:cauchy_kernel}
\end{equation}

Finally, the Kullback-Leibler (KL) divergence is used as a measure of agreement
between distributions $\mathbf{P}$ and $\mathbf{Q}$

\begin{equation}
C = \text{KL}(\mathbf{P} \mid \mid \mathbf{Q}) = \sum_{ij} p_{ij} \log \frac{p_{ij}}{q_{ij}}.
\label{eq:kl_divergence}
\end{equation}

\noindent The t-SNE objective is to find embeddings $\mathbf{Y}$ such that the
KL divergence is minimized.

The corresponding gradient takes the form
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \sum_{j \neq i} \left ( p_{ij} - q_{ij} \right ) \left ( \mathbf{y}_i - \mathbf{y}_j \right ) \left ( 1 + || \mathbf{y}_i - \mathbf{y}_j || ^2 \right )^{-1}.
\label{eq:tsne_gradient}
\end{equation}

Optimization is performed using batch gradient descent with the delta-bar-delta
update rule~\cite{jacobs1988increased}. Originally, t-SNE was run for 1000
iterations consisting of two phases. In the first \textit{early exaggeration}
phase, the attractive forces between data points is increased by a factor
$\rho=12$ so that points can more easily move throughout the embedding to find
their respective neighbors. The remaining 750 iterations are run with $\rho=1$,
which reverts the attractive forces to their original values and produces the
final embedding.

Belkina \textit{et al.} later found that faster convergence can be achieved by
increasing the learning rate to $\eta=N/12$~\cite{belkina2019automated}. The
number of iterations can then safely be lowered to 750, leading to faster
runtime. This convention has been adopted by most modern t-SNE implementations.

\subsection*{Efficient Approximation Schemes} A direct evaluation of t-SNE
gradients requires $\mathcal{O}(N^2)$ operations, which makes its application
impractical to any reasonably-sized data set and beckons for the development of
efficient approximation schemes. Van der Maaten observed that the t-SNE
gradient may be reformulated in terms of a N-body problem where data points may
be interpreted as particles which attract and repel each
other~\cite{van2014accelerating}. The gradient may be rewritten as
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \left [ \sum_{j \neq i} p_{ij} q_{ij} Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) -\sum_{j \neq i} q_{ij}^2 Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) \right ], \label{eq:grad_attr_rep}
\end{equation}
where $Z = \sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2
\right )^{-1}$. The two terms may be viewed as the attractive and repulsive
forces between particles, respectively. Each term lends itself to efficient
approximations, which enable us to greatly reduce the time complexity of t-SNE.

\subsubsection*{Attractive Forces}

Van der Maaten observed that since $p_{ij}$ is calculated in the
high-dimensional space using a Gaussian kernel and that the bandwidth is
selected such that only a fixed number of closest neighbors (determined by the
perplexity parameter) will contribute to the attractive forces of a given point
due to the exponential decay of the Gaussian kernel~\cite{van2014accelerating}.
Therefore it is sufficient to calculate the attractive forces only over a
relatively small number of nearest neighbors instead of all $N$ points. By
utilizing tree-based nearest-neighbor search methods, the time complexity is
thus reduced to $\mathcal{O}(N \log N)$. Linderman \textit{et al.} further
realized that, qualitatively, embeddings are visually virtually
indistinguishable when using only \textit{approximate} nearest neighbors,
further reducing time complexity to $\mathcal{O}(N)$~\cite{linderman2019fast}.

\subsubsection*{Repulsive Forces}

Examining the second term of Eqn.~(\ref{eq:grad_attr_rep}) we notice that each
point indiscriminately exerts a repulsive force on all other points. Van der
Maaten proposed an approach based on N-body simulations and used a
space-partitioning Barnes-Hut tree approach to approximate the interaction
between data points~\cite{van2014accelerating}. Briefly, the approach splits
the space into quadrants (in the 2D case) and if a given point is far enough
away from a quadrant, the points in that quadrant are summarized by a single
center of mass. This reduces the time complexity from $\mathcal{O}(N^2)$ to
$\mathcal{O}(N \log N)$.

More recently, Linderman \textit{et al.} proposed an alternative approach,
FIt-SNE, based on non-uniform convolutions to calculating all pairwise
interactions between repelling data points~\cite{linderman2019fast}. Briefly,
Linderman \textit{et al.} observed that the repulsive forces $\mathbf{R}$ may
be rewritten as \begin{align}
\mathbf{R}_i &= \sum_{j \neq i} q_{ij}^2 Z \left ( \mathbf{y}_i - \mathbf{y}_j \right ) \notag \\
&= \sum_{j \neq i} \frac{\mathbf{y}_i - \mathbf{y}_j}{\left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 \right )^{2}}
\bigg/
\sum_{k \neq l} \frac{1}{1 + || \mathbf{y}_k - \mathbf{y}_l || ^2} \label{eq:fitsne_rep}
\end{align} 
and computed by evaluating three terms
\begin{align}
\phi_{1, j} &= \sum_{j \neq i} \frac{1}{1 + || \mathbf{y}_j - \mathbf{y}_i ||^2}, \notag \\
\phi_{2, j} &= \sum_{j \neq i} \frac{\mathbf{y}_j}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 \right)^2}, \notag \\
\phi_{3, j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 \right)^2}. \label{eq:fitsne_terms}
\end{align}
Then the numerator and denominator of Eqn.~(\ref{eq:fitsne_rep}) can be
computed as $\mathbf{y}_i \phi_{1,j} - \phi_{2,j}$ and $Z = \sum_j \phi_{3,j}$,
respectively. These interactions are accelerated by interpolating these terms
through a grid of equispaced interpolation points. This shifts the
computational burden onto the interpolation points and reduces the time
complexity to $\mathcal{O}(N)$.

\opentsne\ provides efficient implementations for all the aforementioned
improvements but defaults to using approximate nearest neighbor search and the
non-uniform convolutions approach as these enable the user to quickly create
informative visualizations for data sets containing up to millions of samples.

\subsection*{Embedding New Samples}

t-SNE is non-parametric and does not define an explicit mapping from the
high-dimensional space to the embedding space. Therefore embeddings of new data
points need to be found through the use of optimization
techniques~\cite{policar2019embedding}. When adding new data points to an
existing, reference embedding, the reference data points are fixed in place and
the new data points are allowed to find their respective positions. The
optimization remains the same as in standard t-SNE with only slight
modifications to $p_{ij}$ and $q_{ij}$
\begin{align}
p_{j \mid i} &= \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) /  \sigma_i^2 \right )}{\sum_{i} \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) / \sigma_i^2 \right )}, \\
q_{j \mid i} &= \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}}{\sum_{i}\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}},
\end{align}
\noindent where $\mathbf{V} = \left \{ \mathbf{v}_1, \mathbf{v}_2, \dots,
\mathbf{v}_M \right \} \in \mathbb{R}^D$ where $M$ is the number of samples in
the secondary data set and $\mathbf{W} = \left \{ \mathbf{w}_1, \mathbf{w}_2, \dots,
\mathbf{w}_M \right \} \in \mathbb{R}^d$. Additionally, we omit the
symmetrization step in Eqn.~(\ref{eq:symmetrize}). The gradients of
$\mathbf{w}_j$ with respect to the loss (Eqn.~\ref{eq:kl_divergence}) are:
\begin{equation}
\frac{\partial C}{\partial \mathbf{w}_j} = 2 \sum_i \left ( p_{j \mid i} - q_{j \mid i} \right ) \left ( \mathbf{y}_i - \mathbf{w}_j \right ) \left ( 1 + || \mathbf{y}_i - \mathbf{w}_j || ^2 \right )^{-1}
\label{eq:gradient}
\end{equation}

As in standard t-SNE, a direct calculation of gradients takes $\mathcal{O}(N
\cdot M)$ time, but it is straightforward to adapt the Barnes-Hut and FIt-SNE
approximation schemes, reducing the time complexity to $\mathcal{O}(M \log N)$
and $\mathcal{O}(\max \{ N, M \})$, respectively. In the FIt-SNE approximation
scheme, we additionally exploit the fact that the reference embedding remains
fixed throughout the optimization of newly added points, and precompute the
interpolation grid. This further reduces the runtime complexity from
$\mathcal{O}(\max \{ N, M \})$ to $\mathcal{O}(M)$.

\subsection*{Alternative Perplexity Kernels}

In standard t-SNE, distances are converted to similarities through the use of
Gaussian kernels of varying bandwidths. The bandwidths are indirectly
determined by the user-specified perplexity parameter so that a fixed number of
nearest data points will have non-zero similarity. One common trick for
uncovering the global relations between clusters is to increase perplexity so
that more long-range interactions are preserved in the final embedding.
However, one unfortunate side effect of increasing perplexity is that smaller
clusters get absorbed into larger ones.

Kobak \& Berens~\cite{kobak2019art} suggest that replacing the Gaussian kernel
with a mixture of Gaussians may provide better insight into both the local and
global structure. Therefore, the similarities between data points in the input
space become

\begin{align}
  p_{j\mid i} &\propto \frac{1}{\sigma_{1,i}} \exp \left ( -\mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / 2\sigma_{1,i}^2 \right ) \notag \\
  &+ \frac{1}{\sigma_{2,i}} \exp \left ( -\mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / 2\sigma_{2,i}^2 \right ).
\end{align}

The bandwidths of each Gaussian $\sigma_{1,i}$ and $\sigma_{2,i}$ is determined
by selecting different perplexity values. They show that using a kernel with
perplexity 500 captures the long-range interactions which preserve global
cluster organization in the final embedding and combine this with a kernel with
perplexity 50, which prevents small, well-defined clusters from being absorbed
into larger ones.

\subsection*{Variable Degrees of Freedom}

Standard t-SNE reveals the clustering structure in data at one level of
resolution. Different perplexity parameter values can be used to identify
global cluster relationships or small, well-isolated groups. Unfortunately,
varying perplexity values can be time-consuming as this involves recomputing
the KNNG which is often the most expensive part of the t-SNE algorithm.
Alternatively, Kobak \textit{et al.}~\cite{kobak2019heavy} suggest that varying
the degree of freedom in the t-distribution used in the embedding space during
optimization can be used to explore the clustering structure at different
levels of resolution.

Standard t-SNE models similarities between data points in the embedding space
using a t-distribution with a single degree of freedom

\begin{equation}
q_{ij} \propto \left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 / \alpha \right )^{-\alpha} = \frac{1}{\left( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 / \alpha \right)^\alpha }.
\end{equation}

In standard t-SNE $\alpha=1$ so this simplifies to the Cauchy kernel from
Eqn.~(\ref{eq:cauchy_kernel}).

The gradient of the loss function then becomes
\begin{align}
\frac{\partial C}{\partial \mathbf{y}_i} &= 4 \sum_{j \neq i} \left ( p_{ij} - q_{ij} \right ) q_{ij}^{1/\alpha} \left ( \mathbf{y}_i - \mathbf{y}_j \right ),
\end{align}
which can, again, be decomposed into the attractive and repulsive forces as
\begin{align}
\frac{\partial C}{\partial \mathbf{y}_i} &= 4 \sum_{j \neq i} p_{ij} q_{ij}^{1/\alpha} (\mathbf{y}_i - \mathbf{y}_j) \notag \\
&- 4 \sum_{j \neq i} q_{ij}^{\frac{\alpha+1}{\alpha}} / Z (\mathbf{y}_i - \mathbf{y}_j).
\end{align}

Adapting existing approximation schemes to this formulation is straightforward.
We provide efficient implementations of both the Barnes-Hut and the FIt-SNE
approximation schemes, where we replace the terms from
Eqn.~(\ref{eq:fitsne_terms}) with

\begin{align}
\phi_{1,j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)^{\alpha+1}}, \notag \\
\phi_{2,j} &= \sum_{j \neq i} \frac{\mathbf{y}_j}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)^{\alpha+1}}, \notag \\
\phi_{3,j} &= \sum_{j \neq i} \frac{1}{\left( 1 + || \mathbf{y}_j - \mathbf{y}_i ||^2 / \alpha \right)}. \notag
\end{align}

\subsection*{Variable Exaggeration}

Embeddings produced by standard t-SNE often produce clusters separated by thin
boundaries and make use of all available space. While this is often desirable,
this often obscures the global relationships between clusters as all
neighboring clusters appear at the same distance from one another. Other
dimensionality reduction methods such as UMAP~\cite{2018arXivUMAP} or
ForceAtlas2~\cite{jacomy2014forceatlas2} tend to produce embeddings where
clusters appear more compact and the white-space separating the clusters may be
interpreted at least partially as a loose measure of distance.

Kobak \textit{et al.}~\cite{todo} showed that the early exaggeration factor
$\rho$ in t-SNE may be increased to produce more compact layouts often seen in
UMAP or ForceAtlas2. Early exaggeration is a trick introduced by van der
Maaten~\cite{maaten2008visualizing} used in the first 250 iterations and
increases the attractive forces between data points, enabling neighboring
points to move close to one another while ignoring the repulsive forces between
data points. In standard t-SNE, the remainder of the iterations are run with
$\rho=1$. Kobak \textit{et al.} showed that using $\rho=4$ and $\rho=30$ in the
second phase of the optimization results in embeddings visually similar to UMAP
embeddings, and ForceAtlas2 embeddings, respectively.

While it is difficult to claim one is better than the other, different settings
of $\rho$ may uncover different properties and clustering structures. For
example, in single-cell data, standard t-SNE with $\rho=1$ often uncovers
distinct groups of cell-types but obscures developmental paths. This problem is
exacerbated when dealing with large numbers of data points. Conversely,
ForceAtlas2 has often been used to uncover trajectories and transitions between
cell types, but larger clusters often absorb small, distinct groups of cells.

\subsection*{Globally Consistent Initialization Schemes}

Similarly to UMAP and ForceAtlas2, t-SNE can also be viewed as a force-directed
layout algorithm. The first phase of the t-SNE algorithm constructs a
$k$-nearest neighbor graph, and the subsequent optimization balances the
attractive and repulsive forces between data points.

The final positioning of data points using this class of algorithms is largely
dependent on the initial positions in the embedding space. UMAP is often said
to better preserve global structure than t-SNE, but Kobak \& Linderman have
recently discovered that this supposed advantage is entirely due to the better
default initialization in UMAP than in older implementations of
t-SNE~\cite{kobak2019umap}. Standard implementations of t-SNE use random
initializations, which often scatter resulting clusters throughout the
embedding space. UMAP injects global organization into the embedding by
initializing point positions using Laplacian eigenmaps. When initialized in the
same manner, both t-SNE and UMAP perform equally well, or equally poorly,
dispelling the myth that UMAP is superior in this respect.

\opentsne\ defaults to using the two leading principal components as
initialization, but also provides a spectral approach, similar to that of UMAP.
The clusters in resulting embeddings are organized in a globally coherent
manner, leading to more interpretable visualizations.

\section*{Availability}

openTSNE is distributed under the BSD-3-Clause License and is publicly
available as an open-source package at
\url{https://github.com/pavlin-policar/openTSNE}. openTSNE is also available on
PyPI and conda-forge. The data sets used and scripts used in this study are
included in accompanying notebooks, publicly available at
\url{https://github.com/pavlin-policar/opentsne-paper}.

\begin{backmatter}

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Author's contributions}
P.G.P. implemented and maintains \opentsne library. M.S. helped in development of mathematical foundations and advised on inclusion of the extensions. B.Z. initiated and supervised the project. All three co-authors wrote the manuscript.

\section*{Acknowledgements}
We want to thank Dmitry Kobak for various helpful discussions and best practices when using t-SNE and his contributions to the source code. We would also like to thank George Linderman for his help with the FIt-SNE algorithm.

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{references}      % Bibliography file (usually '*.bib' )

\end{backmatter}
\end{document}