### 2022-04-08 paper/main.tex
Finish up implementation section

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

% Custom imports, and defns
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\opentsne}{\pkg{openTSNE}}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{
  Pavlin G. Poli\v{c}ar\\
  Faculty of Computer and Information Science\\
  University of Ljubljana\\
  Ljubljana, Slovenia
  \AND
  Martin Stra\v{z}ar\\
  Broad Institute of MIT and Harvard\\
  Cambridge, MA, USA
  \AND
  Bla\v{z} Zupan\\
  Faculty of Computer and Information Science\\
  University of Ljubljana\\
  Ljubljana, Slovenia
}
\Plainauthor{P. G. Poli\v{c}ar, M. Stra\v{z}ar, B. Zupan}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding}
\Plaintitle{openTSNE: a modular Python library for t-SNE dimensionality reduction and embedding}
\Shorttitle{openTSNE: Extensible, parallel implementations of t-SNE}

%% - \Abstract{} almost as usual
\Abstract{
One of the most popular techniques for constructing visualizations of large, high-dimensional data sets is $t$-distributed stochastic neighbor embedding (t-SNE). Recently, several extensions have been proposed to address scalability issues and the quality of the resulting visualizations. We introduce \opentsne, a modular Python library that implements the core t-SNE algorithm and its many extensions. The library is fast and can treat data sets containing millions of data points. \opentsne\ is open-source and is distributed under the BSD-3-Clause License. Its source code is available at \url{https://github.com/pavlin-policar/openTSNE}.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{t-SNE, embedding, visualization, dimensionality reduction, \proglang{Python}}
\Plainkeywords{t-SNE, embedding, visualization, dimensionality reduction}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Pavlin G. Poli\v{c}ar\\
  Faculty of Computer and Information Science, University of Ljubljana\\
  Ve\v{c}na pot 113, Ljubljana, Slovenia\\
  E-mail: \email{pavlin.policar@fri.uni-lj.si}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction]{Introduction} \label{sec:intro}

The ever-growing volumes of high-dimensional data sets in machine learning call for efficient dimensionality reduction techniques and their implementations, capable of producing informative data visualizations. Popular approaches include principal component analysis, multidimensional scaling, $t$-distributed stochastic neighbor embedding (t-SNE)~\citep{maaten2008visualizing}, and uniform manifold approximation and projections (UMAP)~\citep{2018arXivUMAP}. Among these, t-SNE has received much attention as it can address high volumes of data and reveal the underlying data structure. For instance, t-SNE is widely used in the bioinformatics community in areas such as single-cell transcriptomics~\citep{macosko2015highly,cao2019single,tasic2018shared}, human genetics~\citep{hirata2019genetic}, metagenomic assembly~\citep{beaulaurier2018metagenomic}, the spatial organization of microbial communities~\citep{sheth2019spatial}, and metabolomics~\citep{tkachev2019differences}. To visually explore the data structure, the objects of interest, such as diseased and healthy tissues or single cells, are profiled through thousands of features that include, for example, expression of genes or concentration of metabolites. The task of t-SNE is to embed the objects of interest within a low, preferably two-dimensional embedding. Reports on single-cell gene expression data, our running example, often start with an overview of the cell landscape and use t-SNE to embed high-dimensional expression profiles into a two-dimensional space. Figs.~\ref{fig:macosko}.a and \ref{fig:macosko}.b show two such embeddings.

\begin{figure*}[ht]
  \includegraphics[width=\textwidth]{macosko2015}
  \caption{\label{fig:macosko}
We use \opentsne\ to generate three t-SNE embeddings and demonstrate recent theoretical advances. The data in \textbf{(a)} and \textbf{(b)} includes 44,808 mouse retinal cells (dots in the plots) that are described with high-dimensional gene-expression profiles from \citet{macosko2015highly}. The data in \textbf{(c)} additionally contains 27,499 expression profiles from mouse retinal cells from \citet{shekhar2016comprehensive}. \textbf{(a)} We construct t-SNE embedding following the parameter choices from the original publication by \citet{maaten2008visualizing}. The visualization shows no preservation of the global organization of clusters, resulting from random initialization and an affinity model focused on preserving local neighborhoods. \textbf{(b)} A modern t-SNE embedding, utilizing the latest theoretical advances and practical recommendations constructed using a multi-scale affinity model, preserving both short-range and long-range interactions between data points and initialized so that the global layout is as meaningful as possible. Unlike in \textbf{(a)}, the green and blue clusters representing different sub-types of amacrine and bipolar cells are now localized to the same regions of the space, indicating a higher level of similarity than to other cell types. The embedding in \textbf{(c)} shows how existing t-SNE reference atlases can be used to place new samples into existing embeddings. The positions of new data points correspond to cell types from the reference atlas.
}
\end{figure*}

Despite its utility, t-SNE has often been criticized for its limited scalability, lack of global organization -- t-SNE identifies well-defined clusters that may be arbitrarily scattered throughout the embedding, and the absence of theoretically-founded methods to map new data into existing embeddings~\citep{ding2018interpretable,becht2019dimensionality}. Most of these shortcomings have recently been addressed. \citet{linderman2019fast} developed FIt-SNE, an efficient approximation scheme that massively improves the scalability of t-SNE, achieving linear time complexity in the number of samples. \citet{kobak2019art} proposed several techniques to improve global cluster coherence, including estimating similarities with a mixture of Gaussian kernels. In our previous work, we introduced a principled approach for embedding new samples into existing visualizations~\citep{policar2021embedding}.

However, easy access to efficient implementations almost universally correlates with the widespread adoption of novel computational techniques. Despite the many theoretical advancements, most popular t-SNE libraries have been slow to incorporate them into their implementations. In this paper, we present \opentsne, our open-source Python implementation of t-SNE, complete with its recently proposed extensions. \opentsne\ is easy to install and includes precompiled binaries available through popular Python package managers. It provides a familiar API, making it suitable as a drop-in replacement for existing t-SNE implementations.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Methods} \label{sec:methods}

We first introduce the relevant notation and briefly review the core t-SNE
algorithm in Sec.~\ref{sec:meth.tsne}. We then address the criticisms of
scalability (Sec.~\ref{sec:meth.approx}), the ability to add new samples to
existing embeddings (Sec.~\ref{sec:meth.transform}), and improvements that
improve the global consistency of resulting embeddings
(Sec.~\ref{sec:meth.global}).

\subsection{$t$-distributed stochastic neighbor embedding} \label{sec:meth.tsne}

$t$-distributed stochastic neighbor embedding (t-SNE) is a non-linear
dimensionality reduction method that to aims to find a low-dimensional embedding
where neighborhoods are preserved. More formally, given a multi-dimensional data set
$\mathbf{X} = \left \{ \mathbf{x}_1, \mathbf{x}_2, \dots, \mathbf{x}_N \right \}
\in \mathbb{R}^D$ where $N$ is the number of data points in the data set, t-SNE
aims to find a low dimensional embedding $\mathbf{Y} = \left \{ \mathbf{y}_1,
\mathbf{y}_2, \dots, \mathbf{y}_N \right\} \in \mathbb{R}^d$ where $d \ll D$,
such that if points $\mathbf{x}_i$ and $\mathbf{x}_j$ are close in the
high-dimensional space, their corresponding embeddings $\mathbf{y}_i$ and
$\mathbf{y}_j$ are also close. Since t-SNE is primarily used as a visualization
tool, $d$ is typically set to two. The similarity between two data points in the
high-dimensional space is defined using the Gaussian kernel
\begin{equation}
p_{j \mid i} = \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_j ) / \sigma_i^2 \right )}
{\sum_{k \neq i } \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{x}_k ) / \sigma_i^2 \right )}, \quad p_{i \mid i} = 0
\label{eq:p_ij}
\end{equation}
where $\mathcal{D}$ is some distance measure. The values $p_{j \mid i}$
are then symmetrized to
\begin{equation}
p_{ij} = \frac{p_{j \mid i} + p_{i \mid j}}{2N}.
\label{eq:symmetrize}
\end{equation}

The bandwidth $\sigma_i$ of each Gaussian kernel is selected such that the perplexity of the distribution matches a user-specified parameter value
\begin{equation}
\text{Perplexity} = 2^{H(P_i)} \notag
\end{equation}
where $H(P_i)$ is the Shannon entropy of $P_i$,
\begin{equation}
H(P_i) = -\sum_i p_{j \mid i} \log_2 (p_{j \mid i}). \notag
\end{equation}
This enables t-SNE to adapt to the varying density of the data in the
multi-dimensional space. The perplexity can be interpreted as the continuous
analogue to the number of nearest neighbors to which the distances will be
preserved.

The similarity between points $\mathbf{y}_i$ and $\mathbf{y}_j$ in the embedding
space is defined using the $t$-distribution with a single degree of freedom
(Cauchy kernel)
\begin{equation}
q_{ij} = \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{y}_j ||^2 \right )^{-1}}
{\sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2 \right )^{-1}},
\quad q_{ii} = 0.
\label{eq:cauchy_kernel}
\end{equation}

We use the Kullback-Leibler (KL) divergence to measure the agreement
between the two distributions $\mathbf{P}$ and $\mathbf{Q}$
\begin{equation}
C = \text{KL}(\mathbf{P} \mid \mid \mathbf{Q}) = \sum_{ij} p_{ij} \log \frac{p_{ij}}{q_{ij}}.
\label{eq:kl_divergence}
\end{equation}
The objective is to find embeddings $\mathbf{Y}$ that minimize the KL divergence. The corresponding gradient takes the form
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \sum_{j \neq i} \left ( p_{ij} - q_{ij} \right ) \left ( \mathbf{y}_i - \mathbf{y}_j \right ) w_{ij},
\label{eq:tsne_gradient}
\end{equation}
where $w_{ij} = \left ( 1 + || \mathbf{y}_i - \mathbf{y}_j || ^2 \right )^{-1}$
and represents the unnormalized $q_{ij}$.

Optimization is performed using batch gradient descent using the delta-bar-delta update rule~\citep{jacobs1988increased}. The t-SNE optimization procedure consists of two phases: in the first \textit{early exaggeration} phase, the attractive forces between data points are increased by some factor $\rho$, typically set to $12$, so that points in the embedding can more easily move throughout the space and settle near their respective neighbors. In the second phase of the optimization, the attractive forces are then reverted to their original values with $\rho=1$.

\citet{belkina2019automated} later found that convergence can be sped up by
increasing the learning rate from the standard $\eta=200$ to $\eta=N/12$. As a
side-effect, embeddings converge faster, and the number of iterations can be
lowered from the typical 1000 to 750, decreasing the overall runtime.

\subsection{Efficient Approximation Schemes} \label{sec:meth.approx}

A direct evaluation of t-SNE gradients requires $\mathcal{O}(N^2)$ operations,
which makes its application impractical to any reasonably-sized data set and
highlights the need for the development of efficient approximation schemes.
\citet{van2014accelerating} observed that the t-SNE gradient can be cast as an
N-body problem where data points represent particles that attract and repel each
other. The gradient from Eqn.~(\ref{eq:tsne_gradient}) can be rewritten as
\begin{equation}
\frac{\partial C}{\partial \mathbf{y}_i} = 4 \Bigg [
\underbrace{\sum_{j \neq i} p_{ij} q_{ij} Z \left ( \mathbf{y}_i - \mathbf{y}_j \right )}_{\text{attractive forces}}  -
\underbrace{\sum_{j \neq i} q_{ij}^2 Z \left ( \mathbf{y}_i - \mathbf{y}_j \right )}_{\text{repulsive forces}}
\Bigg ], \label{eq:grad_attr_rep} \notag
\end{equation}
where $Z = \sum_{k \neq l}\left ( 1 + || \mathbf{y}_k - \mathbf{y}_l ||^2 \right
)^{-1}$. We can view this equation as a particle simulation, where the
two terms represent the attractive and repulsive forces between individual
particles. Each term lends itself to efficient approximations, enabling us to
reduce the time complexity of t-SNE to $\mathcal{O}(N)$. This allows modern
t-SNE implementations to be leveraged for the visualization of data sets
containing up to millions of data points.

\subsubsection*{Attractive Forces}
\citet{van2014accelerating} observed that evaluating the attractive forces
between all pairs of data points is excessive, and that considering only a
handful of nearest neighbors at each point are sufficient to obtain a good
approximation. Therefore, instead of calculating all pairwise interactions, it
is sufficient to only find and evaluate the attraction to a certain number of
$k$ nearest neighbors. Using tree-based nearest-neighbor search methods,
\citet{van2014accelerating} reduced the time complexity to $\mathcal{O}(N \log
N)$. \citet{linderman2019fast} further realized that, qualitatively, embeddings
are visually indistinguishable when using only \textit{approximate} nearest
neighbors, further reducing time complexity to $\mathcal{O}(N)$.

\subsubsection*{Repulsive Forces}
Similarly, the repulsive term can also be approximated, motivated by methods from
particle simulations. \citet{van2014accelerating} proposed an approach based on
N-body simulations and used a space-partitioning Barnes-Hut tree approach to
approximate the interaction between data points. This reduces the time
complexity from $\mathcal{O}(N^2)$ to $\mathcal{O}(N \log N)$. More recently,
\citet{linderman2019fast} proposed an alternative approach, FIt-SNE, based on
non-uniform convolutions, which further reduces the time complexity to
$\mathcal{O}(N)$.

\subsection{Embedding New Samples} \label{sec:meth.transform}

The t-SNE algorithm is non-parametric and does not define an explicit mapping from the
high-dimensional space to the embedding space. Therefore embeddings of new data
points need to be found through optimization
techniques~\citep{policar2021embedding}. When adding new data points to an
existing, reference embedding, the reference data points are fixed in place
while new data points are allowed to find their respective positions. The
optimization remains the same as in standard t-SNE with only slight
modifications to $p_{ij}$ and $q_{ij}$
\begin{align}
p_{j \mid i} = \frac{\exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) /  \sigma_i^2 \right )}{\sum_{i} \exp \left ( -\frac{1}{2} \mathcal{D}(\mathbf{x}_i, \mathbf{v}_j) / \sigma_i^2 \right )}, \qquad
q_{j \mid i} = \frac{\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}}{\sum_{i}\left ( 1 + || \mathbf{y}_i - \mathbf{w}_j ||^2 \right )^{-1}}, \notag
\end{align}
\noindent where $\mathbf{V} = \left \{ \mathbf{v}_1, \mathbf{v}_2, \dots,
\mathbf{v}_M \right \} \in \mathbb{R}^D$ where $M$ is the number of samples in
the new data set and $\mathbf{W} = \left \{ \mathbf{w}_1, \mathbf{w}_2, \dots,
\mathbf{w}_M \right \} \in \mathbb{R}^d$. Additionally, we omit the
symmetrization step in Eqn.~(\ref{eq:symmetrize}). Plugging these terms into
Eqn.~(\ref{eq:kl_divergence}), we obtain the following gradient
\begin{equation}
\frac{\partial C}{\partial \mathbf{w}_j} = 2 \sum_i \left ( p_{j \mid i} - q_{j \mid i} \right ) \left ( \mathbf{y}_i - \mathbf{w}_j \right ) \left ( 1 + || \mathbf{y}_i - \mathbf{w}_j || ^2 \right )^{-1}. \notag
\label{eq:gradient}
\end{equation}

Similarly to standard t-SNE, a direct calculation of gradients takes
$\mathcal{O}(N \cdot M)$ time, but it is straightforward to adapt the Barnes-Hut
and FIt-SNE approximation schemes, reducing the time complexity to
$\mathcal{O}(M \log N)$ and $\mathcal{O}(M)$, respectively. Special care must be
taken to tune the learning rate during optimization as the default values
$\eta=200$ and $\eta=N/12$ result in highly unstable optimization.

\subsection{Improving the Global Structure of Embeddings} \label{sec:meth.global}

One of t-SNE's long-standing criticisms is that it fails to preserve long-range
distances but instead focuses on capturing the local structure of the data
manifold~\citep{becht2019dimensionality}. Recently, several approaches have been
proposed to improve the global organization of the resulting embeddings.

\citet{kobak2019umap} showed that the initialization of force-directed layout
algorithms, such as t-SNE, UMAP~\citep{2018arXivUMAP}, and
ForceAtlas2~\citep{jacomy2014forceatlas2}, largely dictates the global
consistency of resulting embeddings. When initialized randomly, clusters are
often arbitrarily scattered in the embedding space. However, when using
initialization schemes based on PCA or spectral embeddings, the clusters
identified in the resulting embedding are typically grouped in a globally
coherent manner.

In standard t-SNE, distances between data points are converted to similarities
through the use of Gaussian kernels. The perplexity of these kernels governs the
sizes of the local neighborhoods which are to be preserved. One easy way to
improve the global consistency of resulting embeddings is to increase the sizes
of these neighborhoods. This can be achieved by increasing the perplexity
parameter. However, increasing perplexity often leads to loss in local sturcture
and tends to obscure small clusters. \citet{kobak2019art} instead propose to use
mixtures of Gaussians to preserve better short-range and long-range distances
and, thus, achieve a better trade-off between local and global structure.

Embeddings produced by standard t-SNE often use all available space and separate
clusters by only thin boundaries. When working with large data sets, this often
obscures the global relationships between clusters as all neighboring groups
appear equidistant from one another.  Other dimensionality reduction methods,
such as UMAP and ForceAtlas2, produce embeddings where clusters appear more
compact, and the white space separating the clusters may be interpreted as a
loose measure of distance. Recently, \citet{bohm2020unifying} showed that the
exaggeration factor $\rho$ could be used to produce layouts more similar to UMAP
and ForceAtlas2. By incorporating exaggeration into later phases of the
optimization, t-SNE introduces white space between clusters, which may better
reflect the global relations between clusters.

Standard t-SNE reveals the clustering structure at a single level of resolution.
While the perplexity parameter can be used to control the trade-off between
local and global structure, this can be time-consuming, and small, well-defined
clusters can be missed. Alternatively, \citet{kobak2019heavy} suggest that
varying the degrees of freedom in the $t$-distribution can be used to explore
the clustering structure at different levels of resolution. Modifying
Eqn.~(\ref{eq:cauchy_kernel}) to $q_{ij} \propto \left ( 1 + || \mathbf{y}_i -
\mathbf{y}_j ||^2 / \alpha \right )^{-\alpha}$ allows us to use heavier-tailed
distributions to model distances in the embedding space, which acts to highlight
small subgroups in the resulting visualization.

\section{Implementation} \label{sec:implementation}

We introduce \opentsne, a comprehensive Python library that implements the t-SNE algorithm and its numerous recently proposed extensions. \opentsne\ is designed to accommodate all types of users and includes both a simple API that allows users to quickly obtain informative t-SNE visualizations, as well as a more flexible API that allows seasoned practitioners to modify all the
different aspects of the t-SNE algorithm.

The core t-SNE algorithm is comprised of three main steps, mirrored in the
structure of the \opentsne\ library:

\begin{enumerate}
\item Firstly, an affinity model describing the similarities between data points must be specified. Standard t-SNE predominantly deals with tabular data, and similarities are computed using Eqn.~\ref{eq:p_ij}. However, \citet{kobak2019art} showed that using mixtures of Gaussians instead of a single Gaussian often reveals better global consistency of the resulting visualization (albeit at some computational cost). Alternatively, one can speed up the overall computation by choosing a uniform kernel, which yields quantitatively similar results to standard t-SNE. \code{openTSNE.affinity} provides several affinity models to choose from but defaults to the standard perplexity-based model proposed in the original publication.

Furthermore, it is straightforward to define and use a custom affinity model. For instance, one can easily extend the \code{openTSNE.affinity.Affinities} class and provide an affinity model for relational data using a normalized adjacency matrix. This enables rapid experimentation and opens the doors to the visualization of non-tabular data.

\item Secondly, initial positions must be chosen for each data point in the embedding space. These can either be chosen randomly or can be used to impose some initial global structure on the embedding, e.g., using a PCA-based initialization. Accordingly, the \code{openTSNE.initialization} module provides several possible initialization schemes, or the user can also specify their own initialization by providing a \code{numpy.ndarray} object. By default, \opentsne\ opts for PCA-based initialization.

\item Finally, equipped with an affinity model and initial point positions, the only remaining step is to optimize the embedding such that the positions of the points in the embedding space reflect their corresponding similarities in the affinity model. In \opentsne, the initialization and affinities are used to construct an \code{openTSNE.TSNEEmbedding} object, which can then be optimized using different parameter settings. Once an embedding object is constructed, it can subsequently be optimized via the \code{.optimize} method.

Optimization can be performed using either the Barnes-Hut or FIt-SNE approximation scheme. While the FIt-SNE approximation scheme scales linearly with the number of samples, it introduces additional computational overhead, which is often unnecessary for smaller data sets. For smaller data sets, this sometimes results in longer runtimes than when using the Barnes-hut approximation scheme. By default, \opentsne\ uses a heuristic to ensure the fastest algorithm is chosen for the given data set, but this setting can also be specified by the user.
\end{enumerate}

Once the embedding is constructed and optimized, it is ready to use for any downstream tasks. \code{openTSNE.TSNEEmbedding} objects inherit from \code{numpy.ndarray}, which enables easy manipulation and provides interoperability with the broader \proglang{Python} ecosystem.

Embedding new data points to an existing \code{openTSNE.TSNEEmbedding}
follows the same three general steps as constructing a standard embedding:
\begin{enumerate}
\item Firstly, an affinity model must be provided describing the similarities between each new data point and the data points in the reference embedding. As before, \opentsne\ provides several possible affinity models in the \code{openTSNE.affinity} module. Most often, however, we want to utilize the same affinity model that was used to construct the reference embedding, as this leads to more spatially consistent embeddings. This is the default behavior of \opentsne.

\item Secondly, the initial positions of the new data points in the embedding space must be specified. These can, again, be initialized randomly or be informed by their similarity to the data points in the reference embedding. As before, \opentsne\ provides several initialization options in the \code{openTSNE.initialization} module, but the user is free to provide their own initialization.

\item Finally, what remains is to construct an embedding object, which can then be subject to optimization. Because this new embedding relies on an existing reference embedding, we construct a \code{openTSNE.PartialTSNEEmbedding} embedding object, which indicates its dependence on a reference \code{openTSNE.TSNEEmbedding} object. As before, \code{openTSNE.PartialTSNEEmbedding} inherits from \code{numpy.ndarray} and can therefore be easily manipulated. Optimization is, again, performed via the \code{.optimize} method.
\end{enumerate}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.75\textwidth]{opentsne-diagram}
  \caption{\label{fig:diagram}
  The high-level structure and flow of the \opentsne\ library.
  }
\end{figure*}

Fig.~\ref{fig:diagram} shows the high-level structure and flow of the \opentsne\ library.

Through different parameters, the user is given full control over every aspect of the t-SNE visualization algorithm, from different ways to define affinities to the delicate tuning of optimization parameters. This allows for rapid experimentation and enables users to quickly and efficiently inspect their data from numerous angles and at varying levels of resolution.

While \opentsne\ allows for an unprecedented level of control at every step, most end-users often only need a quick and easy way to create informative visualizations of their data sets. \opentsne\ provides a high-level API that closely follows the style of \pkg{scikit-learn}~\citep{sklearn_api}, a popular and widely used toolkit in the machine learning community. This allows users of all levels to quickly pick up and make use of the \opentsne\ library.

\subsection{Example}
\label{sec:iris}

\subsubsection{Installation}

\opentsne\ is designed to be easy to use and can be installed through either \pkg{PyPI} using \pkg{pip}
\begin{CodeChunk}
\begin{CodeInput}
pip install opentsne
\end{CodeInput}
\end{CodeChunk}
or using \pkg{conda} from the \pkg{conda-forge} channel
\begin{CodeChunk}
\begin{CodeInput}
conda install -c conda-forge opentsne
\end{CodeInput}
\end{CodeChunk}
\opentsne\ provides precompiled binaries for all major platforms (Windows, MacOS, Linux) for all currently supported \proglang{Python} versions (3.7---3.10). This is especially convenient for Windows users, as this obviates the need for a \proglang{C}/\proglang{C++} compiler, which is typically not bundled with the operating system.

To verify that that \opentsne\ was installed correctly, try to import the module:
\begin{CodeChunk}
\begin{CodeInput}
>>> import openTSNE
>>> print(openTSNE.__version__)
0.6.2
\end{CodeInput}
\end{CodeChunk}

\subsubsection{A simple workflow}

We demonstrate the simplest usage of \opentsne\ on the Iris data set from \citet{anderson1936species}. This small data set describes 150 iris flowers containing four variables (sepal length, sepal width, petal length, petal width) and their species (Iris Setosa, Iris Versicolor, Iris Virginica), which is the target variable. This data set is included in \pkg{scikit-learn}:
\begin{CodeChunk}
\begin{CodeInput}
>>> from sklearn import datasets
>>>
>>> iris = datasets.load_iris()
>>> iris.data[:3]
array([[5.1, 3.5, 1.4, 0.2],
       [4.9, 3. , 1.4, 0.2],
       [4.7, 3.2, 1.3, 0.2]])
\end{CodeInput}
\end{CodeChunk}
\pkg{scikit-learn} loads the data set into a \pkg{numpy} array, which can directly
be used by \opentsne:
\begin{CodeChunk}
\begin{CodeInput}
>>> import openTSNE
>>>
>>> embedding = openTSNE.TSNE().fit(iris.data)
>>> embedding[:3]
TSNEEmbedding([[-0.55581917, 17.56162344],
               [-1.76742536, 20.07763453],
               [-0.58444608, 20.00052495]])
\end{CodeInput}
\end{CodeChunk}
Fig.~\ref{fig:iris} shows a visualization of the resulting t-SNE visualization. As expected, t-SNE is correctly able to group together samples from the same species. Additionally, the t-SNE embedding reveals that Iris Setosa is markedly different from the other two species, while flowers from Iris Versicolor and Iris Virginica may appear more similar to one another.
\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{iris}
  \caption{\label{fig:iris}
  A visualization of the obtained t-SNE embedding on the Iris data set. The colors
  correspond to different species of the Iris flower.
}
\end{figure*}
We can now use this embedding as a reference for new data points. In this simple example, we will duplicate some of the original samples and add them into the embedding.
\begin{CodeChunk}
\begin{CodeInput}
>>> new_data = iris.data[::3]
>>> new_embedding = embedding.transform(new_data)
>>> new_embedding[:3]
PartialTSNEEmbedding([[-0.02637324, 15.69956674],
                      [-2.84200792, 16.15478451],
                      [-2.4594136 , 15.37910876]])
\end{CodeInput}
\end{CodeChunk}
Fig.~\ref{fig:iris-transform} shows the resulting positions of the new data points, overlaid the original, reference embedding. As expected, t-SNE is correctly able to place the new data points to their corresponding groups.
\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{iris-transform}
  \caption{\label{fig:iris-transform}
  A visualization of the original, reference data set (transparent points) overlaid with the new data points (opaque points).
}
\end{figure*}

\subsubsection{Peeking under the hood}

While the above example offers users a quick and easy way to generate informative visualizations, \opentsne\ allows us to control the t-SNE algorithm at every step. To demonstrate the flexibility of \opentsne, we will repeat the same analysis of the Iris data set as before, this time leveraging the more advanced features of the library. More practical examples of advanced usage will be shown in Section~\ref{sec:discussion}.

We, again, begin by loading the Iris data set using \pkg{scikit-learn}:
\begin{CodeChunk}
\begin{CodeInput}
>>> from sklearn import datasets
>>>
>>> iris = datasets.load_iris()
>>> iris.data[:3]
array([[5.1, 3.5, 1.4, 0.2],
       [4.9, 3. , 1.4, 0.2],
       [4.7, 3.2, 1.3, 0.2]])
\end{CodeInput}
\end{CodeChunk}
Next, we import the \opentsne\ library:
\begin{CodeChunk}
\begin{CodeInput}
>>> import openTSNE
\end{CodeInput}
\end{CodeChunk}
We now mirror the three steps listed above. First, we select an affinity model which describes the similarities between data points:
\begin{CodeChunk}
\begin{CodeInput}
>>> affinities = openTSNE.affinity.PerplexityBasedNN(iris.data, perplexity=30)
\end{CodeInput}
\end{CodeChunk}
Next, we determine the initial point positions in the embedding:
\begin{CodeChunk}
\begin{CodeInput}
>>> initialization = openTSNE.initialization.pca(iris.data)
>>> initialization[:3]
array([[-1.30971087e-04,  1.55848907e-05],
       [-1.32435711e-04, -8.63672049e-06],
       [-1.40967409e-04, -7.07276279e-06]])
\end{CodeInput}
\end{CodeChunk}
Next, using the affinity model and initialization, we can create an embedding object:
\begin{CodeChunk}
\begin{CodeInput}
>>> embedding = openTSNE.TSNEEmbedding(initialization, affinities)
>>> embedding[:3]
TSNEEmbedding([[-1.30971087e-04,  1.55848907e-05],
               [-1.32435711e-04, -8.63672049e-06],
               [-1.40967409e-04, -7.07276279e-06]])
\end{CodeInput}
\end{CodeChunk}
Notice that the embedding object contains the same point coordinates as the initialization. This is because although we have created an \code{openTSNE.TSNEEmbedding} object, we have not yet run any optimization. To optimize the embedding, we merely have to call the \code{embedding.optimize} method using our desired parameter settings. Here, we use the default parameter settings, including a shorter early-exaggeration phase, followed by a longer phase with no exaggeration.
\begin{CodeChunk}
\begin{CodeInput}
>>> embedding.optimize(250, exaggeration=12, momentum=0.5, inplace=True)
>>> embedding.optimize(500, momentum=0.8, inplace=True)
>>> embedding[:3]
TSNEEmbedding([[-0.55581917, 17.56162344],
               [-1.76742536, 20.07763453],
               [-0.58444608, 20.00052495]])
\end{CodeInput}
\end{CodeChunk}
We can validate that method produces the exact same embedding as before:
\begin{CodeChunk}
\begin{CodeInput}
>>> np.linalg.norm(openTSNE.TSNE().fit(iris.data) - embedding)
0.0
\end{CodeInput}
\end{CodeChunk}
Next, we can add new samples into the embedding:
\begin{CodeChunk}
\begin{CodeInput}
>>> new_embedding = embedding.prepare_partial(iris.data[::3], perplexity=5)
>>> new_embedding.optimize(
...     exaggeration=1.5, n_iter=250, learning_rate=0.1, momentum=0.8,
...     max_grad_norm=0.25, inplace=True
... )
>>> new_embedding[::3]
PartialTSNEEmbedding([[-0.02637324, 15.69956674],
                      [-2.84200792, 16.15478451],
                      [-2.4594136 , 15.37910876]])
\end{CodeInput}
\end{CodeChunk}
And validate that this also produces the exact same embedding as before:
\begin{CodeChunk}
\begin{CodeInput}
>>> np.linalg.norm(new_embedding - embedding.transform(iris.data[::3]))
0.0
\end{CodeInput}
\end{CodeChunk}

\section{Discussion and Examples} \label{sec:discussion}

\subsection{Uncovering Structure in High-Dimensional Data}

Dimensionality reduction techniques implicitly assume that high-dimensional data
lies on a lower-dimensional manifold, which a small number of dimensions can
accurately capture. However, there is no evidence that every data set can
accurately be described using only two dimensions, and any such embedding will
inevitably lead to a loss of information. Thus, it is beneficial to examine
multiple embeddings, each of which provides a different perspective on topology
and other data characteristics.

\begin{figure*}[htbp]
  \center
  \includegraphics[width=0.75\textwidth]{tasic2018}
  \caption{\label{fig:tasic}
  We use \opentsne\ to create four different
  visualizations of the \citet{tasic2018shared} data,
  each providing a different perspective into the topology of the data.
  The data set contains 21,874 single-cells originating from the mouse
  neocortex. Cluster annotations and colors are taken from the original
  publication. Warm colors correspond to excitatory neurons, cool colors
  correspond to inhibitory neurons, and gray/brown colors correspond to
  non-neuronal cells. Standard t-SNE \textbf{(a)} emphasizes local
  structure while increasing perplexity \textbf{(b)} results in a more
  meaningful layout of the clusters. We can also combine the two
  perplexities by using a multi-scale kernel affinity model \textbf{(c)}
  and obtain a trade-off between global and local structure.
  Alternatively, we can inspect more fine-grained structure and reveal
  smaller clusters by using a more heavy-tailed kernel \textbf{(d)}.}
\end{figure*}

We illustrate this point by generating four different embeddings of the data on
single-cell gene expression in mouse brain~\citep{tasic2018shared}.
Fig.~\ref{fig:tasic}.a shows an embedding using default t-SNE parameters. We generate
the four different embeddings using the following code snippet (shortened for clarity).
\begin{CodeChunk}
\begin{CodeInput}
embedding_a = openTSNE.TSNE(perplexity=30).fit(x)
embedding_b = openTSNE.TSNE(perplexity=500).fit(x)
embedding_c = openTSNE.TSNE(perplexity=[30, 500]).fit(x)
embedding_d = openTSNE.TSNE(perplexity=30, dof=0.6).fit(x)
\end{CodeInput}
\end{CodeChunk}
While different clusters of excitatory and inhibitory neurons appear close to one another, all clusters appear equidistant from their neighbors, and the overall relations between groups are not obvious. The embedding in Fig.~\ref{fig:tasic}.b focuses on preserving larger neighborhoods of points, resulting in a more globally consistent layout where relations between clusters become more apparent. Here, it is evident from the increased white space between groups that there is one large class of excitatory neurons and two related classes of inhibitory neurons. Unfortunately, focusing on preserving large neighborhoods leads to the absorption of smaller clusters into larger ones. Alternatively, Fig.~\ref{fig:tasic}.c uses multi-scale similarity kernels that aim to preserve both the global organization of clusters and prevent smaller cluster absorption. We constructed the embedding from Fig.~\ref{fig:tasic}.d with the settings used for Fig.~\ref{fig:tasic}.a, but at a finer level of resolution. The figure demonstrates that some clusters are composed of numerous, smaller subgroups representing different cell subpopulations that are not visible under standard parameter settings.

\subsection{Embedding Massive Data Sets}

When dealing with data containing millions of data points, standard t-SNE
embeddings often become unwieldy -- cluster boundaries are blurred, large
clusters absorb smaller ones, and relationships between clusters become
increasingly difficult to interpret. We constructed Fig.~\ref{fig:cao}.a from
the data containing expression profiles of over two million single cells
captured at different time points in mouse development. The embedding indicates
numerous clusters with transitions between time points, as marked by the
color-coding, that are difficult to interpret. \citet{kobak2019art} that
increasing attractive forces between similar data points controlled via
the \textit{exaggeration} parameter leads to more compact clusters, and
subsequently, more informative visualizations. For instance, \ref{fig:cao}.b
doubles the default exaggeration, which uncovers some
of the data's overall structure. Further doubling the exaggeration in
\ref{fig:cao}.c allows us to observe that the data is comprised of two main groups
of cells and eight somewhat smaller clusters. The visualization also reveals
several tiny clusters, possibly corresponding to rare cell types.
\begin{CodeChunk}
\begin{CodeInput}
# To improve the global structure of the final layout, we initialize the
# embedding by creating a t-SNE plot on a sample of points

# Take a random sample of the data points
indices = np.random.permutation(list(range(x.shape[0])))
reverse = np.argsort(indices)

x_sample, x_rest = x[indices[:25000]], x[indices[25000:]]

# Create a t-SNE embedding on the sample, using a high perplexity value to
# emphasize the high level structure
init_sample = openTSNE.TSNE(perplexity=500).fit(x_sample)

# We can next use the `transform` functionality to determine starting positions
# for the remaining data points. Since we don't actually want to optimize the
# remaining points separately from the sample points, we don't need to call the
# `transform` function. Instead, we'll call the `prepare_partial` function,
# which `transform` calls anyways under the hood.
init_rest = init_sample.prepare_partial(x_rest)

# Lastly, we merge `init_sample` and `init_rest` and restore the original order
init_full = np.vstack((init_sample, init_rest))[reverse]
\end{CodeInput}
\end{CodeChunk}
We generate the three different embeddings using the following code snippet
(shortened for clarity).
\begin{CodeChunk}
\begin{CodeInput}
# We separately precompute the affinities as this can take quite a long time.
# This way, we can reuse the precomputed affinities in various different
# embeddings, without having to repeat the same, costly computation every time.
affinities = openTSNE.affinity.PerplexityBasedNN(x, perplexity=30)

# Create the TSNEEmbedding object using our previously calculated initialization
# and precomputed affinities
embedding = openTSNE.TSNEEmbedding(init_full, affinities)

# Run the standard t-SNE early exaggeration phase
embedding_ee =    embedding.optimize(n_iter=500, exaggeration=12, momentum=0.5)
embedding_c  = embedding_ee.optimize(n_iter=500, exaggeration=4,  momentum=0.8)
embedding_b  =  embedding_c.optimize(n_iter=500, exaggeration=2,  momentum=0.8)
embedding_a  =  embedding_b.optimize(n_iter=500, exaggeration=1,  momentum=0.8)
\end{CodeInput}
\end{CodeChunk}

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{cao2019}
  \caption{\label{fig:cao}
  Increasing the exaggeration parameter leads to compact clusters, highlighting
  the data's global organization and emphasizing the continuous nature of
  cell state transitions. The data set from \citet{cao2019single}
  contains expression profiles from 2,058,652 single cells. The data were
  collected from mice embryos at different developmental stages at daily
  intervals after 9.5 to 13.5 days. \textbf{(c)} reveals that the data is
  comprised of two main components -- the neural tube and mesenchymal
  cells -- as well as several other smaller clusters. The
  colors indicate developmental progression with red indicating
  least-developed cells and blue indicating most developed cells. The
  overall developmental trajectory is most apparent with higher
  exaggeration levels, showing red cells slowly transitioning into blue
  cells. Progressively easing the exaggeration factor uncovers finer
  clusters within the larger groups, as shown in \textbf{(b)} with
  exaggeration of two and subsequently in \textbf{(a)}, where we show the
  standard t-SNE with no exaggeration. 32,011 putative doublets are
  excluded from the visualizations.
}
\end{figure*}

Exaggeration can highlight transitions between cell states in developmental
studies. Standard t-SNE often produces embeddings with clearly defined, discrete
clusters. We can adjust the level of granularity and resolution of the clusters
with several parameters in Fig.~\ref{fig:tasic}. However, discrete clusters are
often undesired in developmental studies where cells' state is assumed to follow
a continuous transition path. To this end, other embedding techniques such as
UMAP and ForceAtlas2 are used to better capture the continuity between cell
states. Recently, \citet{bohm2020unifying} showed that
embeddings produced by t-SNE with exaggeration values of 4 and $\sim30$
construct embeddings that are markedly similar to UMAP and ForceAtlas2,
respectively. For example, in Fig.~\ref{fig:cao}.a, the developmental trajectory
between different time points is difficult to observe due to many sprawled out
clusters. On the other hand, it is easier to trace the development when we
increase the exaggeration factor from $1$ to $2$ to $4$ in
Figs.~\ref{fig:cao}.b-c.

\subsection{Embedding New Samples}

Unlike other popular dimensionality reduction techniques such as principal
component analysis or autoencoders, t-SNE is a non-parametric method and does
not define an explicit mapping to the embedding space. Therefore, embeddings of
new data points need to be found through
optimization~\citep{policar2021embedding}. \opentsne\ is currently the only
publicly available library allowing users to add new samples to existing
embeddings in a principled manner.

Figs.~\ref{fig:macosko}.b and \ref{fig:macosko}.c demonstrate how we can use a
previously labeled single-cell data set and embed cells from a separate
experiment into the reference landscape. The reference data from
\citet{macosko2015highly} contains gene expression profiles from mouse retinal
cells. By embedding the samples from a similar experiment on bipolar retinal
cells by \citet{shekhar2016comprehensive}, we can correctly map the bipolar
cell clusters onto the reference embedding. The following snippet shows the
code necessary to generate figure~\ref{fig:macosko}.
\begin{CodeChunk}
\begin{CodeInput}
# We specify the old default parameter values which are often used to show
# why t-SNE misbehaves or fails to preserve global structure
tsne_old = openTSNE.TSNE(
    initialization="random", learning_rate=200, n_iter=750
).fit(x)

# By default, openTSNE uses PCA-based initialization and automatically
# determines the best learning rate. This also allows us to lower the default
# number of iterations to 500, lowering the overall runtime
tsne_standard = openTSNE.TSNE(perplexity=[50, 500]).fit(x)

# Use the generated embeddings to add new data points
new_embedding = tsne_standard.transform(x_new)
\end{CodeInput}
\end{CodeChunk}

Embedding single cells into existing reference atlases can also be useful for
cell-type classification in cases of unknown cell identities. For instance, in
Fig.~\ref{fig:transform}, we construct a reference embedding using labeled data
from \citet{hochgerner2018conserved} containing gene-expression profiles of
cells from the mouse brain. The authors assign a type to each cell. We can
verify their classification accuracy by visualizing the expression of
well-established gene markers for the major cell types. We then embed cells
from \citet{harris2018classes} into the constructed cell atlas. In
\citet{harris2018classes}, labels are provided only for neuronal cells. In the
resulting mapping, we can quickly identify other non-neuronal cell types,
including oligodendrocytes and astrocytes. We can further use marker genes to
validate that the mapping in the reference landscape is correct.

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{transform_hochgerner}
  \caption{\label{fig:transform}
  \opentsne\ supports embedding new samples into an existing reference t-SNE
  landscape. For the series of visualizations shown in this figure, we
  first construct a t-SNE embedding for the data from 
  \citet{hochgerner2018conserved} containing 24,185
  developing, single cells from the mouse hippocampus. The data contains
  gene expression in different neurons, supporting glia, and other
  vascular cells (upper left). Data points representing cells are colored
  according to cell-types assigned in the original publication; see the
  legend from Fig.~\ref{fig:tasic} to map colors to cell-type. We then
  embed new, hippocampal cells collected in a study by
  \citet{harris2018classes} using the embedding of \citet{hochgerner2018conserved}
  data as a reference. In their study, \citet{harris2018classes}
  collected 6,971 single-cells and focused on identifying different
  types of inhibitory neurons. However, almost half of the collected cells
  are not neurons and were left uncharacterized. Inspecting the embeddings
  of these cells in the reference embedding (bottom left) reveals that in
  addition to inhibitory neurons, the data contains several supporting
  glial cells as well as a small population of endothelial cells. We can
  verify our approach's accuracy by inspecting marker genes for the major
  cell types in the reference (top row) and embedded samples (bottom row).
}
\end{figure*}

The examples presented above demonstrate how to use \opentsne\ to quickly gain
insight into newly-sequenced, single-cell data sets by utilizing existing cell
atlases. The approach is general and not limited to single-cell gene expression,
and one can, in principle, apply it to any tabular data set regardless of field.

\subsection{Versatility}

Versatility, the ability to use and combine different optimization approaches to
construct different embedding spaces, is another of \opentsne's core design
principles. \citet{kobak2019art} recently provided several recommendations and
tricks to obtain better and more meaningful t-SNE visualizations. These include
multi-scale similarity kernels, perplexity annealing, and increasing
exaggeration when working with massive data sets. \opentsne\ provides a flexible
API to incorporate these improvements in just a few lines of code. Furthermore,
\opentsne\ supports custom affinity models, enabling users to construct t-SNE
embeddings on non-tabular relational data: the only requirement imposed by the
affinity model is some notion of similarity between data points. Finally,
\opentsne's comprehensive callback system can be utilized to monitor and adapt
different stages of the optimization phase and has been used to construct
visually appealing animations of the t-SNE optimization process.

\subsection{Speed}

% General problem statement
One of the t-SNE's common criticisms is limited scalability to large data sets containing, for instance, millions of data points~\citep {becht2019dimensionality}. The culprit for slow response time stems from particular optimization procedures and their specific implementation in popular open-source libraries. Until quite recently, most popular implementations of t-SNE were based on either a direct implementation of the t-SNE algorithm with asymptotic time complexity $\mathcal{O}(N^2)$~\citep{maaten2008visualizing}, or the Barnes-Hut approximation scheme developed by \citet{van2014accelerating} with asymptotic time complexity $\mathcal{O}(N \log N)$. Recently, however, \citet{linderman2019fast} developed a novel approximation scheme -- FIt-SNE -- which reduces the asymptotic time complexity to $\mathcal {O}(N)$, overcoming the difficulties of scaling t-SNE to large data sets.

% We benchmark Python
Fig.~\ref{fig:benchmarks_py} benchmarks four popular Python~\citep{vanrossum1995python} t-SNE
implementations, including \pkg{scikit-learn} (v1.0.1), \pkg{MulticoreTSNE}
(v0.1), \pkg{FIt-SNE} (v1.1.0), and \opentsne\ (v0.6.2). Benchmarks were
performed on an Intel(R) Xeon(R) CPU E5-1650 equipped with 128 GB of memory.
Benchmarks were run for $1,000$ iterations with the original t-SNE parameters,
as some implementations do not allow for their modification. To ensure a
realistic benchmarking scenario, we utilize the 10X Genomics 1.3 million mouse
brain data set~\citep{xxx}, containing 1.3 million single-cell gene-expression
profiles. We preprocess the data using the standard single-cell analysis
pipeline~\citep{xxx} and extract the top 50 principal components. To generate
benchmark data sets of different sample sizes, we subsample the data set five
times at each sample size. In total, we run each implementation on 30 different
data sets.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.75\textwidth]{benchmarks_python-final}
  \caption{\label{fig:benchmarks_py}
  We benchmark \opentsne\ (v0.6.2) against three popular open-source
  implementations from \pkg{scikit-learn}~\citep{pedregosa2011scikit}
  (v1.0.1), \pkg{MulticoreTSNE}~\citep{Ulyanov2016} (v0.1), and
  \pkg{FIt-SNE}~\citep{linderman2019fast} (v1.1.0). Experiments were run on a
  on an Intel(R) Xeon(R) CPU E5-1650 equipped with 128 GB of memory. Notice
  that \opentsne\ scales similarly to \pkg{FIt-SNE}, as they both use the
  same interpolation-based approximation scheme, while \pkg{scikit-learn} and
  \pkg{MulticoreTSNE} utilize the Barnes-Hut approximation.
}
\end{figure*}

% Problem with Python
In Python, the most widely-used implementation of t-SNE comes from \pkg{scikit-learn}, which exhibits long runtimes when compared to other Python implementations. \pkg{scikit-learn}, like its \proglang{C++} counterpart -- \pkg{MulticoreTSNE}~\citep{Ulyanov2016} , is based on the Barnes-Hut approximation scheme. However, as shown on Fig.~\ref{fig:benchmarks_py}, \pkg{scikit-learn} exhibits much longer runtimes than \pkg{MulticoreTSNE}. Newer implementations, such as \pkg{FIt-SNE} and \opentsne, both implement the FIt-SNE approximation scheme, making them suitable for the visualization of millions of data points. While both libraries implement the same algorithm, \opentsne\ emphasizes ease of use and extensibility and is primarily written in Python, which inevitably introduces some runtime overhead. This makes \opentsne\ marginally slower than its pure \proglang{C++} counterpart -- \pkg{FIt-SNE}. However, these differences disappear almost entirely when utilizing multi-threading, as shown in Fig.~\ref{fig:benchmarks_py}. Additionally, \opentsne\ provides a flexible API, allowing the user to split up the embedding-construction process into several parts. This allows us to cache and reuse intermediate results, enabling users to quickly experiment with different parameter settings and iterate on their final visualizations. This kind of functionality is not supported in other implementations, which require the user to start from scratch for every embedding, resulting in large amounts of redundant computation.

% Problem with other programming languages
In other programming languages, the selection of fast, open-source implementations of t-SNE is more limited. In \proglang{R}~\citep{team2013r}, the most widely used implementation of t-SNE -- \pkg{Rtsne}~\citep{krijthe2015rtsne} -- is based on the Barnes-Hut approximation, and is closely comparable to \pkg{MulticoreTSNE}. In \proglang{Julia}~\citep{bezanson2017julia}, the most popular implementation -- \pkg{TSne.jl}~\citep{julia_tsne} -- is based on a naive $\mathcal{O}(N^2)$ implementation of the algorithm, which allows it to be used only on smaller data sets containing up to thousands of data points. We benchmark \pkg{Rtsne} (v0.15) and  \pkg{TSne.jl} (v1.3.0) and compare it to \opentsne\ (v0.6.2). Fig.~\ref{fig:benchmarks_lang} illustrates the scaling of each approximation scheme. Of these implementations, only \opentsne\ is able to scale to data sets containing millions of data points in a reasonable amount of time.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.75\textwidth]{benchmarks_langs-final}
  %\includegraphics[width=\textwidth]{benchmarks-combined}
  \caption{\label{fig:benchmarks_lang}
  We select the fastest available, open-source implementations of the t-SNE
  algorithm from the \proglang{Python}, \proglang{R}, and \proglang{Julia}
  programming languages. We choose
  \opentsne\ (v0.6.2) for Python, \pkg{Rtsne}~\citep{krijthe2015rtsne} (v0.15)
  for \proglang{R}, and \pkg{TSne.jl}~\citep{julia_tsne} (v1.3.0) for
  \proglang{Julia}. The benchmarks reflect the time complexity of the
  implemented approximation schemes. For instance, \pkg{TSne.jl} implements a
  naive $\mathcal{O}(N^2)$ algorithm, making it prohibitively expensive for all
  but the smallest of data sets. \pkg{Rtsne} scales similarily to
  \pkg{MulticoreTSNE} in Python, as it implements the $\mathcal{O}(N \log N)$
  Barnes-Hut approximation scheme. Interestingly, the multi-threaded
  implementation version of \pkg{Rtsne} crashes on benchmark data sets
  containing over 500,000 data points due to lack of stack memory. We found
  these crashes to be inconsistent and somewhat random. \opentsne\ is the only
  library implementing the more recent FIt-SNE approximation scheme, which makes
  it suitable for larger data sets.
}
\end{figure*}

\subsection{Ease of Use}

Intuitive access and simple installation procedures almost universally correlate
with the widespread adoption of novel computational techniques. While the t-SNE
implementation from \pkg{scikit-learn} fits this requirement, the implementation
proves prohibitively slow for even moderately-sized data sets that span tens of
thousands of data records. Other \proglang{C++} implementations such as \pkg{MulticoreTSNE}
and \pkg{FIt-SNE} exhibit better scaling in more massive data sets, but do not
provide precompiled binaries and require users to compile the software
themselves. This problem is critical, for instance, for users of the Windows
operating system, where the \proglang{C++} compiler does not come with the system, making
the correct configuration of current t-SNE implementations cumbersome.

We designed \opentsne\ to be accessible to a broader audience. We provide
precompiled binaries for all major Python versions on all major platforms,
making the installation process as seamless as possible. One can install
\opentsne\ through the Python Package Index (\pkg{PyPI}) or \pkg{conda} from the
\pkg{conda-forge} channel, the two most widely adopted Python package managers.
\opentsne's high-level interface is inspired by \pkg{scikit-learn}, which is well
established in the Python data science ecosystem. \opentsne\ implements
multi-threaded versions of both the Barnes-Hut and FIt-SNE approximation
schemes, enabling it to be applied to data sets containing millions of data
points. While the Python virtual machine inevitably introduces some performance
overhead, the runtime is comparable to its \proglang{C++} counterpart -- \pkg{FIt-SNE}.
Finally, \opentsne\ is extensible. Its modular design enables researchers to
quickly experiment with different parameter settings and easily incorporate
custom components into the software. We provide an extended list of features and
comparison to other popular t-SNE implementations in Tab.~\ref{tab:features}.

\begin{table}
\begin{center}\small
\newcommand*\rot{\rotatebox{90}}
\renewcommand{\arraystretch}{1.25}

\begin{tabular}{l c c c c|c c}
\toprule
\setlength\tabcolsep{6pt}
& \rot{\pkg{scikit-learn}} & \rot{\pkg{MulticoreTSNE}} & \rot{\pkg{FIt-SNE}} & \rot{\pkg{openTSNE}} & \rot{\pkg{Rtsne}} & \rot{\pkg{Tsne.jl}} \\

\toprule
\textbf{Ease of Installation} \\
Native Python installation & \checkmark & & & \checkmark & & \\
\pkg{PyPI} package & \checkmark & \checkmark & & \checkmark & & \\
\pkg{conda} package & \checkmark & & & \checkmark & & \\

\hline
\textbf{Approximation Schemes} \\
None ($\mathcal{O}(N^2)$) & & & & & & \checkmark \\
Barnes-Hut ($\mathcal{O}(N \log N)$) & \checkmark & \checkmark & & \checkmark & \checkmark & \\
FIt-SNE ($\mathcal{O}(N)$) & & & \checkmark & \checkmark & & \\

\hline
\textbf{Advanced Features and Extensions} \\
%Multiscale Gaussian kernels & & & & \checkmark & \checkmark & & \\
Extensible affinity models & & & & \checkmark & & \\
Variable degrees of freedom & & & \checkmark & \checkmark & & \\
Variable exaggeration & & & \checkmark & \checkmark & & \\
Globally-consistent initialization & & & \checkmark & \checkmark & & \\
Automatic learning rate & & & \checkmark & \checkmark & & \\
Adding samples to embeddings & & & & \checkmark & & \\
Callback system & & & & \checkmark & & \\
Interactive optimization  & & & & \checkmark & & \\

\hline
\textbf{Project Quality}\\
Actively maintained & \checkmark & & & \checkmark & \checkmark & \checkmark \\
Continuous integration & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark \\
User documentation & \checkmark & & & \checkmark & \checkmark  & \\
End-to-end usage examples & \checkmark & & \checkmark & \checkmark & \checkmark & \checkmark \\
API reference & \checkmark & & & \checkmark & \checkmark &  \\
\bottomrule
\end{tabular}
\end{center}

\caption{\label{tab:features}
  We compare the features of \opentsne~(v0.6.2) to \pkg{scikit-learn}~(v1.0.1),
  \pkg{BHTSNE}~(master), \pkg{MulticoreTSNE}~(v0.1), and
  \pkg{FIt-SNE}~(v1.1.0), as well as the major implementations in \proglang{R}
  and \proglang{Julia}, \pkg{Rtsne}~(v0.15), \pkg{Tsne.jl}~(v1.3.0). Active
  maintenance indicates whether the project has had any meaningful development
  in the past year from the time of writing.
}
\end{table}

\section{Conclusion}

We introduce \opentsne, the most feature-complete, open-source, Python
implementation of the widely popular t-SNE visualization algorithm. The library
incorporates the latest developments to the t-SNE algorithm and makes them
easily accessible through a familiar API. \opentsne\ implements efficient
approximation schemes, making it suitable for the visualization of large data
sets containing up to millions of data points. Finally, \opentsne\ is easy to
install and comes with precompiled binaries available for all major platforms.

%% -- Optional special unnumbered sections -------------------------------------

\section*{Availability and Documentation}

\opentsne\ is distributed under the BSD-3-Clause License and its source code is
publicly available at \url{https://github.com/pavlin-policar/openTSNE}.
\opentsne\ can also be installed from \pkg{PyPI} and \pkg{conda-forge}.
End-to-end usage examples, documentation, and API reference are all available at
\url{https://opentsne.readthedocs.io}.

\section*{Reproducibility}

The results in this paper were obtained using \proglang{Python}~3.10.2 with the
\pkg{openTSNE}~v0.6.2 package. All results and figures can be reproduced using
the scripts found in the accompanying GitHub repository
\url{https://github.com/pavlin-policar/opentsne-paper}.

\section*{Acknowledgments}

We want to thank Dmitry Kobak for helpful discussions regarding t-SNE and his
contributions to the source code. We would also like to thank George Linderman
for his help with the FIt-SNE algorithm. This work was supported by Slovenian
Research Agency (P2-0209).

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{references}

\end{document}